---
title: "Soil p, k, pH"
author: "All of us"
format:
  html:
    code-fold: false
    embed-resources: true
    toc: true
    number-sections: true
    theme: journal
    #cerulean, journal, quartz, sketchy, slate, morph, superhero, vapor
knitr: 
  opts_chunk: 
    message: false
    warning: false
---

# Learning Objectives  

Work on Soil P, K and pH data, and interpolate the interpolated map for these variables 

# Setup  
```{r setup }
# Packages
library(dplyr)
library(tidyr)
library(janitor)
library(readr)
library(sf) #vector manipulation
library(ggplot2)
library(viridis)
library(ggthemes)
library(gstat) #IDW
library(stars) #raster manipulation
library(terra) #topographical variables
library(patchwork) #combining multiple plots  
library(tibble) #rownames to column
library(ggspatial)
```
Define map theme for re-use
```{r defining map theme}
mytheme_map <- 
  theme_map()+
  theme(legend.position = "bottom",
        legend.direction = "vertical",
        panel.background = element_rect(fill="white",
                                        color=NA),
        plot.background = element_rect(fill="white",
                                       color=NA),
        legend.title = element_text(size=10),
        legend.text = element_text(size=9),
        plot.margin = margin(5, 10, 5, 10))

```

```{r reading soil-grid data}
soil_grid = read_sf("../data/50ac/soilgrid/BackFieldPivot_SampleData_2025-02-12.shp")%>%
  clean_names()

soil_grid
summary(soil_grid)

```

- select column needed (id, bufferph, k, p)
- rename columns to include units if needed
- set crs
- transform crs from geometric to projected

# Wrangling  
```{r grided soil data}
soil_grid_w = soil_grid %>%
  select(id, 
         ph = bufferph,
         p_lbac = p,
         k_lbac = k) %>%
  st_set_crs(value = 4326) %>%
  st_transform(crs = 6345)

soil_grid_w 
```

# EDA  
```{r ph plot}
soil_grid_w %>%
  ggplot() +
  geom_sf(aes(color = ph),
          size = .5) +
  scale_color_viridis_b() +
  mytheme_map
```


```{r P plot}
soil_grid_w %>%
  ggplot() +
  geom_sf(aes(color = p_lbac),
          size = .5) +
  scale_color_viridis_b() +
  mytheme_map

```


```{r K plot}
soil_grid_w %>%
  ggplot() +
  geom_sf(aes(color = k_lbac),
          size = .5) +
  scale_color_viridis_b() +
  mytheme_map

```

# Grid  
Since we are interpolating, we need to create a grid (same as for the yield and terrain exercises) based on the field boundary.    

```{r boundary}

# bring in the field boundary too
boundary = read_sf("../data/50ac/boundary/BackFieldPivot.shp")%>%
  clean_names() %>%
  st_transform(crs = 6345)

boundary
```
creating the grid needed for interpolating soil variables
```{r grid_r}
# grid in vector format
grid_r = boundary %>%
  st_make_grid(cellsize = 5) %>%
  st_as_sf() %>%
  st_rasterize(dx = 5,
               dy = 5) %>%
  st_crop(boundary)

#grid_r =  grid_r[boundary]

ggplot()+
  geom_stars(data = grid_r)+
  mytheme_map

```

# Interpolation
## soil pH  
```{r idw_mod}
ph_idw_mod = gstat(formula = ph ~ 1,
                      data = soil_grid_w)
ph_idw_mod


plbac_idw_mod = gstat(formula = p_lbac ~ 1,
                      data = soil_grid_w)
plbac_idw_mod


klbac_idw_mod = gstat(formula = k_lbac ~ 1,
                      data = soil_grid_w)
klbac_idw_mod
```

```{r idw_pred}
ph_idw_pred  = predict(ph_idw_mod,
                          grid_r)
ph_idw_pred


plbac_idw_pred  = predict(plbac_idw_mod,
                          grid_r)
plbac_idw_pred


klbac_idw_pred  = predict(klbac_idw_mod,
                          grid_r)
klbac_idw_pred
```

# Grid extraction  
```{r soil_ph_v}
ph_v = ph_idw_pred %>%
  st_as_sf(as_points = F,
           merge = F) %>%
  dplyr::select(ph = var1.pred) %>%
  mutate(ph_class = case_when(
    ph <= 5.7 ~ "Inhibitory (≤ 5.7)",
    ph > 5.7 & ph < 6 ~ "Suboptimal (5.7–6.0)",
    ph >= 6 & ph <= 6.5  ~ "Optimal (6.0–6.5)",
    ph > 6.5 ~ "Above Optimal (> 6.5)")) %>%
  mutate(ph_class = factor(ph_class,
                           levels = c("Inhibitory (≤ 5.7)", "Suboptimal (5.7–6.0)", "Optimal (6.0–6.5)", "Above Optimal (> 6.5)")))

ph_v
```


```{r plbac_v}
plbac_v <- plbac_idw_pred %>%
  st_as_sf(as_points = F, 
           merge = F) %>%
  dplyr::select(p_lbac = var1.pred) %>%
  mutate(plbac_class = case_when(
    p_lbac <= 30 ~ "Low (0–30)",
    p_lbac > 30 & p_lbac <= 60 ~ "Medium (31–60)",
    p_lbac > 60 & p_lbac <= 100  ~ "High (61–100)",
    p_lbac > 100 ~ "Very High (> 100)")) %>%
  mutate(plbac_class = factor(plbac_class,
                              levels = c("Low (0–30)", "Medium (31–60)", "High (61–100)", "Very High (> 100)")))


plbac_v
summary(plbac_v)
```


```{r klbac_v}
klbac_v <- klbac_idw_pred %>%
  st_as_sf(as_points = F, 
           merge = F) %>%
  dplyr::select(k_lbac = var1.pred) %>%
  mutate(klbac_class = case_when(
    k_lbac <= 60 ~ "Low (0–60)",
    k_lbac > 60 & k_lbac <= 150 ~ "Medium (61–150)",
    k_lbac > 150 & k_lbac <= 250  ~ "High (151–250)",
    k_lbac > 250 ~ "Very High (> 250)")) %>%
  mutate(klbac_class = factor(klbac_class,
                              levels = c("Low (0–60)", "Medium (61–150)", "High (151–250)", "Very High (> 250)")))


klbac_v
summary(klbac_v)
```

# Plotting Soil Variable Maps
```{r ph_map }
ph_map = ggplot() +
  geom_sf(data = ph_v, aes(fill = ph_class),
          color = NA) +
  scale_fill_viridis_d(drop = FALSE,
                       option = "B") +
  geom_sf(data = boundary, fill = NA) +
  labs(fill = "Soil pH",
       title = "Soil pH") +
  mytheme_map +
  annotation_scale(location = "bl",
                   width_hint = 0.3,
                   pad_x = unit(0.1, "in"),
                   pad_y = unit(0.01, "in")) +
  annotation_north_arrow(location = "tl",
                       pad_x = unit(0.1, "in"),
                       pad_y = unit(0.05, "in"),
                       height = unit(0.4, "in"),
                       width = unit(0.4, "in"),
                       style = north_arrow_fancy_orienteering())

ph_map
```


```{r plbac_map }
plbac_map = ggplot() +
  geom_sf(data = plbac_v, aes(fill = plbac_class),
          color = NA) +
  scale_fill_viridis_d(drop = FALSE,
                       option = "B") +
  geom_sf(data = boundary, fill = NA) +
  labs(fill = "P (lb/ac)",
        title = "Phosphorus") +
  mytheme_map

plbac_map
```


```{r klbac_map }
klbac_map = ggplot() +
  geom_sf(data = klbac_v, aes(fill = klbac_class),
          color = NA) +
  scale_fill_viridis_d(drop = FALSE,
                       option = "B") +
  geom_sf(data = boundary, fill = NA) +
  labs(fill = "K (lb/ac)",
        title = "Potassium") +
  mytheme_map

klbac_map
```


# Plotting all together
```{r plotting all together}
ph_map + plbac_map + klbac_map +
  plot_annotation(title = "Interpolated Maps for Soil Variables (50-acre Back Field)",
    theme = theme(
      plot.title = element_text(size = 13, face = "bold", hjust = 0.5)))

ggsave("../output/soilph-p-k.png", 
       width = 12, 
       height = 6)
```



# Merging soil pH, P and K layers  
```{r soil_v }
soil_v = ph_v %>%
  st_join(plbac_v,
          join = st_equals,
          left = T) %>%
  st_join(klbac_v,
          join = st_equals,
          left = T)

soil_v
```

# Exporting  
Exporting to file to use later for management zone creation.  
```{r exporting soil_v}
write_sf(soil_v,
         "../data/soil_v.geojson",
         delete_dsn = T)

```
