---
title: "Soil ECa"
author: "Comfort"
format:
  html:
    code-fold: false
    embed-resources: true
    toc: true
    number-sections: true
    theme: journal
    #cerulean, journal, quartz, sketchy, slate, morph, superhero, vapor
knitr: 
  opts_chunk: 
    message: false
    warning: false
---

# Objectives  

I will import Soil ECa data at two depths, do some interpolation using IDW method make publication ready maps with boundaries and export the soil eca layers to file. 

# Setup  
```{r setup }
# Packages
library(dplyr)
library(tidyr)
library(janitor)
library(readr)
library(sf) #vector manipulation
library(ggplot2)
library(viridis)
library(ggthemes)
library(gstat) #IDW
library(stars) #raster manipulation
library(terra) #topographical variables
library(patchwork) #combining multiple plots  
library(tibble) #rownames to column

```

```{r defining map theme}
mytheme_map <- 
  theme_map()+
  theme(legend.position = "bottom",
        legend.direction = "vertical",
        panel.background = element_rect(fill="white",
                                        color=NA),
        plot.background = element_rect(fill="white",
                                       color=NA),
        legend.title = element_text(size=10),
        legend.text = element_text(size=9),
        plot.margin = margin(5, 10, 5, 10))

```

```{r reading eca data}
eca = read_sf("../data/50ac/ec/VSEC_50_acre.shp") %>%
  clean_names()

eca
summary(eca)
```

```{r boundary_w}
boundary_w = read_sf("../data/50ac/boundary/BackFieldPivot.shp") %>%
  st_transform(crs = 6345)

boundary_w
```


# Wrangling  
```{r eca_w}
eca_w = eca %>%
  dplyr::select(x = longitude,
         y = latitude,
         eca30_dsm = e_cshallow,
         eca90_dsm = e_cdeep,
         elev_ft = elevation) %>%
  st_as_sf(coords = c("x", "y")) %>%
  st_transform(crs = 6345) %>%
  st_filter(boundary_w) 

eca_w 
summary(eca_w)
```

# EDA  
```{r eca30 plot}
eca_w %>%
  ggplot() +
  geom_sf(aes(color = eca30_dsm),
          size = .5) +
  scale_color_viridis_b() +
  geom_sf(data = boundary_w, fill = NA,
          linewidth = 1)+
  mytheme_map
```


```{r eca90 plot}
eca_w %>%
  ggplot() +
  geom_sf(aes(color = eca90_dsm),
          size = .5) +
  scale_color_viridis_b() +
  geom_sf(data = boundary_w, fill = NA,
          linewidth = 1)+
  mytheme_map

```

# Grid  
Since we are interpolating, we need to create a grid (same as for the yield and terrain exercises) based on the field boundary.    


```{r grid_r}
# grid in vector format
grid_r = boundary_w %>%
  st_make_grid(cellsize = 5) %>%
  st_as_sf() %>%
  st_rasterize(dx = 5,
               dy = 5) %>%
  st_crop(boundary_w)

grid_r

ggplot()+
  geom_stars(data = grid_r)+
  mytheme_map

```

# Interpolation
## eca30  
```{r eca30_idw_mod}
eca30_idw_mod = gstat(formula = eca30_dsm ~ 1,
                      data = eca_w)

eca30_idw_mod
```

```{r eca30_idw_pred}
eca30_idw_pred  = predict(eca30_idw_mod,
                          grid_r)

eca30_idw_pred
```

```{r eca30_map }
eca30_map = ggplot() +
  geom_stars(data = eca30_idw_pred) +
  scale_fill_viridis_c(na.value = "white") +
  geom_sf(data = boundary_w, fill = NA,
          linewidth = 1) +
  labs(title = "ECa (shallow)",
       fill = "Eca \n0-30 cm \n(dS/m)") +
  mytheme_map +
  annotation_scale(location = "bl",
                   width_hint = 0.3,
                   pad_x = unit(0.1, "in"),
                   pad_y = unit(0.01, "in")) +
  annotation_north_arrow(location = "tl",
                       pad_x = unit(0.1, "in"),
                       pad_y = unit(0.05, "in"),
                       height = unit(0.4, "in"),
                       width = unit(0.4, "in"),
                       style = north_arrow_fancy_orienteering())
  

eca30_map
```

## eca90  
```{r eca90_idw_mod}
eca90_idw_mod = gstat(formula = eca90_dsm ~ 1,
                      data = eca_w)

eca90_idw_mod
```

```{r eca90_idw_pred}
eca90_idw_pred  = predict(eca90_idw_mod, grid_r)

eca90_idw_pred
```

```{r eca90_map }
eca90_map <- ggplot()+
  geom_stars(data = eca90_idw_pred)+
  scale_fill_viridis_c(na.value = "white") +
  geom_sf(data = boundary_w, fill = NA,
          linewidth = 1) +
  labs(title = "ECa (shallow)",
       fill="ECa\n0-90 cm\n(dS/m)")+
  mytheme_map

eca90_map
```

## Plotting all together
```{r plotting all together}
eca30_map + eca90_map + 
  plot_layout(ncol = 2) +
  plot_annotation(title = "Interpolated Electrical Conductivity (ECa) Maps",
    theme = theme(
      plot.title = element_text(size = 13, face = "bold", hjust = 0.5)))

ggsave("../output/soileca.png",
       width = 12,
       height = 4)


```

# Grid extraction  
```{r eca30_v}
eca30_v = eca30_idw_pred %>%
  st_as_sf(as_points = F,
           merge = F) %>%
  dplyr::select(eca30_dsm = var1.pred)

eca30_v
```


```{r eca90_v}
eca90_v <- eca90_idw_pred %>%
  st_as_sf(as_points = F, 
           merge = F) %>%
  dplyr::select(eca90_dsm=var1.pred) 

eca90_v
```

# Merging eca layers  
```{r eca_v }
eca_v = eca30_v %>%
  st_join(eca90_v,
          join = st_equals,
          left = T)

eca_v
```

# Exporting  
Exporting to file to use later for management zone creation.  
```{r exporting}
write_sf(eca_v,
         "../data/eca_v.geojson",
         delete_dsn = T)

```

