---
title: "Yield monitor wrangling for corn, soybeans and cotton"
author: "Comfort Adegbenro"
format:
  html:
    code-fold: false
    embed-resources: true
    toc: true
    number-sections: true
    theme: journal
    #cerulean, journal, quartz, sketchy, slate, morph, superhero, vapor
knitr: 
  opts_chunk: 
    message: false
    warning: false
---

# Learning objectives  
Import yield data for corn, soybeans, and cotton for four different years, do EDA, and then wrangle to get a clean data.

# Setup  
## Loading packages  
```{r setup }
#| message: false

#loading packages (add to the list as we need) 
library(readr)
library(janitor)
library(dplyr)
library(tidyr)
library(ggplot2)
library(sf)
library(viridis)
library(mapview)
library(patchwork)
library(gt)
```

## Importing data  
```{r y16 corn }
files = list.files("../data/50ac/yield/2016 corn/", 
                    pattern = "\\.shp$", 
                    full.names = TRUE) # to import all the file path of each corn yield as a list

y16_corn = lapply(files, read_sf) # this applies the read_sf function to the list of file paths

y16_corn

names(y16_corn) = tools::file_path_sans_ext(basename(files)) #helps to extract and name each source file which is the actual sf object.

lapply(y16_corn, names) #before bindrows(), some of the column names were different, this helped me check which were different and flow/flow_corn, region were identified

y16_corn <- lapply(y16_corn, function(x) {
  if ("Flow" %in% names(x)) {
    x <- rename(x, Flow_Corn = Flow)
  }
  x
}) #this helped replace column flow with flow_corn. I'm leaving region until later, so it populates with NA

lapply(y16_corn, names) #check column names again.

y16_corn_all = bind_rows(y16_corn, .id = "source_file") %>%
  clean_names()

y16_corn_all

```


```{r y17_soybeans}

y17_soybeans = read_sf("../data/50ac/yield/2017 soybeans/2017_50_acre_soybean.shp") %>%
  clean_names()
y17_soybeans

```

```{r y23/24_cotton}

y23_cotton = read_sf("../data/50ac/yield/2023 cotton/50_acre_2023.shp") %>%
  clean_names()
y23_cotton


y24_cotton = read_sf("../data/50ac/yield/2024 cotton/2024_50acre.shp") %>%
  clean_names()
y24_cotton


```

# EDA  
```{r summary y16_corn_all}
summary(y16_corn_all) 
```


```{r summary y17_soybeans}
summary(y17_soybeans) 
```


```{r summary y23/24_cotton}
summary(y23_cotton) 

summary(y24_cotton) 
```

Looking at the distribution of dry yield and speed for corn 2016 

```{r corn 16 yield distribution - density plot}
ggplot(y16_corn_all, aes(x =dry_yield)) +
  geom_density() +
  geom_rug()

```


```{r corn 16 speed distribution - density plot}
ggplot(y16_corn_all, aes(x = speed))+
  geom_density()+
  geom_rug()

```


Looking at the distribution of dry yield and speed for soybean 2017 

```{r soybeans 17 yield distribution - density plot}
ggplot(y17_soybeans, aes(x =dry_yield)) +
  geom_density() +
  geom_rug()

```

```{r soybeans 17 speed distribution - density plot}
ggplot(y17_soybeans, aes(x = speed))+
  geom_density()+
  geom_rug()

```

Looking at the distribution of dry yield and speed for cotton 2023 and 2024 

```{r cotton 23_24 yield distribution - density plot}
ggplot(y23_cotton, aes(x =dry_yield)) +
  geom_density() +
  geom_rug()


ggplot(y24_cotton, aes(x =dry_yield)) +
  geom_density() +
  geom_rug()
```


```{r cotton 23_24 speed distribution - density plot}
ggplot(y23_cotton, aes(x = speed))+
  geom_density()+
  geom_rug()

ggplot(y24_cotton, aes(x = speed))+
  geom_density()+
  geom_rug()

```


```{r corn 16 yield x speed}
ggplot(y16_corn_all, aes(x=speed, 
                    y=dry_yield)) +
  geom_point() +
  geom_smooth()

```


```{r soybeans 17 yield x speed}
ggplot(y17_soybeans, aes(x=speed, 
                    y=dry_yield)) +
  geom_point() +
  geom_smooth()

```


```{r cotton 23_24 yield x speed}
ggplot(y23_cotton, aes(x=speed, 
                    y=dry_yield)) +
  geom_point() +
  geom_smooth()


ggplot(y24_cotton, aes(x=speed, 
                    y=dry_yield)) +
  geom_point() +
  geom_smooth()

```

We need to clean the data 

- Clean based on speed  
- Clean based on yield  

```{r data needs}
y16_corn_all

y17_soybeans

y23_cotton

y24_cotton
```

From a data wrangling perspective, i will select only needed columns which are dry_yield, speed and elevation, with the source-filename (for corn) and geometry

  - speed  
  - dry_yield  
  - elevation
  - source_file  (for corn)

- rename columns to include measurement units  
- set the crs
- transform from geographic (unprojected) to projected  

# Wrangling  
```{r y16_corn_all}
y16_corn_all_w <- y16_corn_all %>% 
  
  # Selecting desired variables and renaming some of them
  select(source_file,
         speed_mph = speed, 
         yield_buac = dry_yield,
         elev_ft = elevation
         ) %>%

  # Setting native crs (geographic)
  st_set_crs(value = 4326) %>%

  # Transforming from geodraphic to projected CRS
   st_transform(crs = 6345)

y16_corn_all_w

summary(y16_corn_all_w)

```

From the summary and density plots, there seems to outlier for corn yield as the maximum is above 300, and speed too with maximum reaching 17


```{r y17_soybeans}
y17_soybeans_w <- y17_soybeans %>% 
  
  # Selecting desired variables and renaming some of them
  select(speed_mph = speed, 
         yield_buac = dry_yield,
         elev_ft = elevation
         ) %>%

  # Setting native crs (geographic)
  st_set_crs(value = 4326) %>%

  # Transforming from geodraphic to projected CRS
   st_transform(crs = 6345)

y17_soybeans_w

summary(y17_soybeans_w)

```

There seem to be outliers for the 2017 soybean yield (274 bu/ac), and a speed of 12.1 mph, seem far away from the median of 3.8 mph.



```{r y23_cotton}
y23_cotton_w <- y23_cotton %>% 
  
  # Selecting desired variables and renaming some of them
  select(speed_mph = speed, 
         yield_lbac = dry_yield,
         elev_ft = elevation
         ) %>%

  # Setting native crs (geographic)
  st_set_crs(value = 4326) %>%

  # Transforming from geodraphic to projected CRS
   st_transform(crs = 6345)

y23_cotton_w

summary(y23_cotton_w)

```

There is an outlier in cotton 2023 yield (46143 lb/ac)



```{r y24_cotton}
y24_cotton_w <- y24_cotton %>% 
  
  # Selecting desired variables and renaming some of them
  select(speed_mph = speed, 
         yield_lbac = dry_yield,
         elev_ft = elevation
         ) %>%

  # Setting native crs (geographic)
  st_set_crs(value = 4326) %>%

  # Transforming from geodraphic to projected CRS
   st_transform(crs = 6345)

y24_cotton_w

summary(y24_cotton_w)

```


# Maps  
```{r static map for 2016 corn}
y16_corn_all_w %>%
  filter(yield_buac < 300) %>%
  ggplot()+
  geom_sf(aes(color = yield_buac), size = .1)+
  #coord_sf(datum = st_crs(y16_corn_all_w)) + #lets you plot it in metres for projected crs
  scale_color_viridis_b()

```

```{r static map for 2017 soybeans}
y17_soybeans_w %>%
  filter(yield_buac < 50) %>%
  ggplot()+
  geom_sf(aes(color = yield_buac), size = .1)+
  #coord_sf(datum = st_crs(y17_soybeans_w)) + #lets you plot it in metres for projected crs
  scale_color_viridis_b()

```



```{r static map for 2023 cotton}
y23_cotton_w %>%
  filter(yield_lbac < 5000) %>%
  ggplot()+
  geom_sf(aes(color = yield_lbac), size = .1)+
  #coord_sf(datum = st_crs(y23_cotton_w)) + #lets you plot it in metres for projected crs
  scale_color_viridis_b()

```

```{r static map for 2024 cotton}
y23_cotton_w %>%
  filter(yield_lbac < 5000) %>%
  ggplot()+
  geom_sf(aes(color = yield_lbac), size = .1)+
  #coord_sf(datum = st_crs(y24_cotton_w)) + #lets you plot it in metres for projected crs
  scale_color_viridis_b()

```


```{r interactive}
y16_corn_all_w %>%
  mapview(zcol ="yield_buac",
          cex = .1,
          lwd = 0)

```


# Exporting  
```{r exporting wrangled y16 corn}
write_sf(y16_corn_all_w,
         "../data/y16corn_processed.geojson",
         delete_dsn = T)

```


```{r exporting wrangled y17 soybeans}
write_sf(y17_soybeans_w,
         "../data/y17soybeans_processed.geojson",
         delete_dsn = T)

```


```{r exporting wrangled y23 cotton}
write_sf(y23_cotton_w,
         "../data/y23cotton_processed.geojson",
         delete_dsn = T)

```


```{r exporting wrangled y24 cotton}
write_sf(y24_cotton_w,
         "../data/y24cotton_processed.geojson",
         delete_dsn = T)

```


# Creating Summaries
```{r summary data y16 corn}
summary_raw_y16_corn = y16_corn_all_w %>%
  st_drop_geometry() %>%
  summarise(Observations = length(yield_buac),
            Min = min(yield_buac),
            Mean = mean(yield_buac),
            Max = max(yield_buac),
            Std_Dev = sd(yield_buac)
            ) %>%
  mutate(Data = "Raw",
         Crop = "Corn",
         Year = 2016) %>%
  dplyr::select(Data, Year, Crop, everything())

summary_raw_y16_corn
```

```{r summary data y17 soybeans}
summary_raw_y17_soybeans = y17_soybeans_w %>%
  st_drop_geometry() %>%
  summarise(Observations = length(yield_buac),
            Min = min(yield_buac),
            Mean = mean(yield_buac),
            Max = max(yield_buac),
            Std_Dev = sd(yield_buac)
            ) %>%
  mutate(Data = "Raw",
         Crop = "Soybeans",
         Year = 2017) %>%
  dplyr::select(Data, Year, Crop, everything())

summary_raw_y17_soybeans
```

```{r summary data y23 cotton}
summary_raw_y23_cotton = y23_cotton_w %>%
  st_drop_geometry() %>%
  summarise(Observations = length(yield_lbac),
            Min = min(yield_lbac, na.rm = TRUE),
            Mean = mean(yield_lbac, na.rm = TRUE),
            Max = max(yield_lbac, na.rm = TRUE),
            Std_Dev = sd(yield_lbac, na.rm = TRUE)
            ) %>%
  mutate(Data = "Raw",
         Crop = "Cotton",
         Year = 2023) %>%
  dplyr::select(Data, Year, Crop, everything())

summary_raw_y23_cotton
```

```{r summary data y24 cotton}
summary_raw_y24_cotton = y24_cotton_w %>%
  st_drop_geometry() %>%
  summarise(Observations = length(yield_lbac),
            Min = min(yield_lbac, na.rm = TRUE),
            Mean = mean(yield_lbac, na.rm = TRUE),
            Max = max(yield_lbac, na.rm = TRUE),
            Std_Dev = sd(yield_lbac, na.rm = TRUE)
            ) %>%
  mutate(Data = "Raw",
         Crop = "Cotton",
         Year = 2024) %>%
  dplyr::select(Data, Year, Crop, everything())

summary_raw_y24_cotton
```


```{r combine all yield summaries}
all_ys = summary_raw_y16_corn %>%
  bind_rows(summary_raw_y17_soybeans) %>%
  bind_rows(summary_raw_y23_cotton) %>%
  bind_rows(summary_raw_y24_cotton)
  
all_ys

all_ys_table = all_ys %>%
  gt() %>%
  tab_header(title = "Yield Summary: Corn, Soybeans and Cotton") %>%
  fmt_number(columns = c(Mean, Std_Dev, Min, Max), decimals = 1) %>%
  fmt_number(columns = Observations, decimals = 0, use_seps = TRUE) %>%
  tab_source_note(source_note = "Yield units: bu/ac for corn and soybean, and lb/ac for cotton") %>%
  tab_options(
    table.width = pct(90),
    table.font.size = 14,
    data_row.padding = px(8),
    column_labels.font.weight = "bold",
    heading.title.font.size = 18,
    heading.title.font.weight = "bold"
  )

all_ys_table

# Saving to load it in another script
saveRDS(all_ys_table, "../output/all_ys_table.rds")
```



# Creating Density Plots
```{r corn 16 yield distribution - density plot}
dplot_y16 = ggplot(y16_corn_all_w, aes(x =yield_buac)) +
  geom_density(fill = "#558B2F",
               color = "black",
               alpha = 0.6,
               linewidth = 0.5) +
  geom_rug(alpha = 0.3) +
  labs(x = " 2016 Corn Yield (bu/ac)",
       y = "Density") +
  theme_classic(base_size = 12)

dplot_y16
```



```{r soybeans 17 yield distribution - density plot}
dplot_y17 = ggplot(y17_soybeans_w, aes(x =yield_buac)) +
  geom_density(fill = "#B8860B",
               color = "black",
               alpha = 0.6,
               linewidth = 0.5) +
  geom_rug(alpha = 0.3) +
  labs(x = " 2017 Soybeans Yield (bu/ac)",
       y = "Density") +
  theme_classic(base_size = 12)

dplot_y17

```


```{r cotton 23 yield distribution - density plot}
dplot_y23 = ggplot(y23_cotton_w, aes(x =yield_lbac)) +
  geom_density(fill = "#6BAED6",
               color = "black",
               alpha = 0.6,
               linewidth = 0.5) +
  geom_rug(alpha = 0.3) +
  labs(x = " 2023 Cotton Yield (lb/ac)",
       y = "Density") +
  theme_classic(base_size = 12)

dplot_y23

```

```{r cotton 24 yield distribution - density plot}
dplot_y24 = ggplot(y24_cotton_w, aes(x =yield_lbac)) +
  geom_density(fill = "#08519C",
               color = "black",
               alpha = 1,
               linewidth = 0.5) +
  geom_rug(alpha = 0.3) +
  labs(x = " 2024 Cotton Yield (lb/ac)",
       y = "Density") +
  theme_classic(base_size = 12)

dplot_y24

```


```{r combine all density plots}
dplot_y16 + dplot_y17 + dplot_y23 + dplot_y24 +
  plot_annotation(title = "Yield Distribution",
                  theme = theme(
      plot.title = element_text(size = 14, face = "bold", hjust = 0.5))) 


ggsave("../output/yield_densitiy_plots.png", 
       width = 10, 
       height = 6)
```

