---
title: "Yield monitor wrangling for corn, soybeans and cotton"
author: "Comfort Adegbenro"
format:
  html:
    code-fold: false
    embed-resources: true
    toc: true
    number-sections: true
    theme: journal
    #cerulean, journal, quartz, sketchy, slate, morph, superhero, vapor
knitr: 
  opts_chunk: 
    message: false
    warning: false
---

# Learning objectives  
Import yield data for corn, soybeans, and cotton for four different years, do EDA, and then wrangle to get a clean data.

# Setup  
## Loading packages  
```{r setup }
#| message: false

#loading packages (add to the list as we need) 
library(readr)
library(janitor)
library(dplyr)
library(tidyr)
library(ggplot2)
library(sf)
library(viridis)
library(mapview)

```

## Importing data  
```{r yield19 }
files = list.files("../data/50ac/yield/2016 corn/", 
                    pattern = "\\.shp$", 
                    full.names = TRUE) # to import all the file path of each corn yield as a list

y16_corn = lapply(files, read_sf) # this applies the read_sf function to the list of file paths

y16_corn

names(y16_corn) = tools::file_path_sans_ext(basename(files)) #helps to extract and name each source file which is the actual sf object.

lapply(y16_corn, names) #before bindrows(), some of the column names were different, this helped me check which were different and flow/flow_corn, region were identified

y16_corn <- lapply(y16_corn, function(x) {
  if ("Flow" %in% names(x)) {
    x <- rename(x, Flow_Corn = Flow)
  }
  x
}) #this helped replace column flow with flow_corn. I'm leaving region until later, so it populates with NA

lapply(y16_corn, names) #check column names again.

y16_corn_all = bind_rows(y16_corn, .id = "source_file") %>%
  clean_names()

y16_corn_all

```

# EDA  
```{r summary}
summary(y16_corn_all) 
```
Looking at the distribution of dry yield and speed for corn 2016 

```{r corn 16 yield distribution - density plot}
ggplot(y16_corn_all, aes(x =dry_yield)) +
  geom_density() +
  geom_rug()

```

```{r corn 16 speed distribution - density plot}
ggplot(y16_corn_all, aes(x = speed))+
  geom_density()+
  geom_rug()

```


```{r corn 16 yield x speed}
ggplot(y16_corn_all, aes(x=speed, 
                    y=dry_yield)) +
  geom_point() +
  geom_smooth()

```
We need to clean the data 

- Clean based on speed  
- Clean based on yield  

```{r data needs}
y16_corn_all
```

From a data wrangling perspective, i will select only needed columns which are dry_yield and speed with the source-filename and geometry

  - speed  
  - dry_yield  
  - source_file  

- rename columns to include measurement units  
- set the crs
- transform from geographic (unprojected) to projected  

# Wrangling  
```{r y16_corn_all}
y16_corn_all_w <- y16_corn_all %>% 
  
  # Selecting desired variables and renaming some of them
  select(speed_mph = speed, 
         yield_buac = dry_yield,
         source_file
         ) %>%

  # Setting native crs (geographic)
  st_set_crs(value = 4326) %>%

  # Transforming from geodraphic to projected CRS
   st_transform(crs = 6345)

y16_corn_all_w

summary(y16_corn_all_w)

```
From the summary and density plots, there seems to outlier for yield as the maximum is above 300, and speed too with maximum reaching 17
# Maps  
```{r static}
y16_corn_all_w %>%
  ggplot()+
  geom_sf(aes(color = yield_buac), size = .1)+
  #coord_sf(datum = st_crs(y16_corn_all_w)) + #lets you plot it in metres for projected crs
  scale_color_viridis_b()

```

```{r interactive}
y16_corn_all_w %>%
  mapview(zcol ="yield_buac",
          cex = .1,
          lwd = 0)

```


# Exporting  
```{r}
write_sf(y16_corn_all_w,
         "../data/y16corn_processed.geojson",
         delete_dsn = T)

```


