---
title: "Yield monitor editor - for corn, soybeans and cotton 2016, 2017, 2023, and 2024"
author: "Comfort Adegbenro"
format:
  html:
    #cerulean, journal, quartz, sketchy, slate, morph, superhero, vapor
    self-contained: true
    code-fold: false
    embed-resources: true
    toc: true
    number-sections: true
    theme: sketchy
knitr: 
  opts_chunk: 
    message: false
    warning: false
editor_options: 
  chunk_output_type: inline
---

# Learning objectives

Clean wrangled yield data through different methods, based on IQR of yields (min and max_y), speed (min and max_v), buffer (removing edge effects)
# Setup

## Loading packages

```{r  }
#| message: false

library(readr)
library(janitor)
library(dplyr)
library(tidyr)
library(ggplot2)
library(sf)
library(viridis)
library(mapview)
library(leafsync)
library(patchwork)
library(ggthemes)
library(ggspatial)
```

## Define map theme for re-use
```{r defining map theme}
mytheme_map <- 
  theme_map()+
  theme(legend.position = "bottom",
        legend.direction = "vertical",
        panel.background = element_rect(fill="white",
                                        color=NA),
        plot.background = element_rect(fill="white",
                                       color=NA),
        legend.title = element_text(size=10),
        legend.text = element_text(size=9),
        plot.margin = margin(5, 10, 5, 10))

```


## Importing data

```{r import all yield data for all crops and boundary}
y16_corn_w = read_sf("../data/y16corn_processed.geojson")
y16_corn_w

y17_soybeans_w = read_sf("../data/y17soybeans_processed.geojson")
y17_soybeans_w

y23_cotton_w = read_sf("../data/y23cotton_processed.geojson")
y23_cotton_w

y24_cotton_w = read_sf("../data/y24cotton_processed.geojson")
y24_cotton_w

boundary = read_sf("../data/50ac/boundary/BackFieldPivot.shp") %>%
  clean_names() %>%
  st_set_crs(4326) %>%
  st_transform(crs = 6345)

boundary
```

# EDA

## Summary

```{r summaries of yields for all crops }
summary(y16_corn_w)
summary(y17_soybeans_w)
summary(y23_cotton_w)
summary(y24_cotton_w)
```

## Raw map

```{r raw maps for corn}
ymap16_corn =  y16_corn_w %>%
  filter(yield_buac < 300) %>%
  ggplot()+
  geom_sf(aes(color = yield_buac),
          size = 0.3) +
  scale_color_viridis_b() +
  geom_sf(data = boundary, fill = NA) +
  labs(color = "Corn yield (bu/acre)",
       title = "Map of Corn Raw Yield Data (outlier of >300 bu/ac excluded)") +
  mytheme_map +
  annotation_scale(location = "bl",
                   width_hint = 0.3,
                   pad_x = unit(0.1, "in"),
                   pad_y = unit(0.01, "in")) +
  annotation_north_arrow(location = "tl",
                       pad_x = unit(0.1, "in"),
                       pad_y = unit(0.05, "in"),
                       height = unit(0.4, "in"),
                       width = unit(0.4, "in"),
                       style = north_arrow_fancy_orienteering())
ymap16_corn
```

# Yield editor - Min/max yield

## Raw yield quantiles

```{r raw yield quantile corn 2016}
lowq_y16_corn  = quantile(y16_corn_w$yield_buac, 0.1)
lowq_y16_corn

upq_y16_corn = quantile (y16_corn_w$yield_buac, 0.99)
upq_y16_corn

iqr_y16_corn = upq_y16_corn - lowq_y16_corn
iqr_y16_corn

miny16_corn = lowq_y16_corn - (0.25*iqr_y16_corn)
miny16_corn

maxy16_corn = upq_y16_corn + (0.25*iqr_y16_corn)
maxy16_corn
```



```{r raw yield quantile soybean 2017}
lowq_y17_soy  = quantile(y17_soybeans_w$yield_buac, 0.1)
lowq_y17_soy

upq_y17_soy = quantile (y17_soybeans_w$yield_buac, 0.99)
upq_y17_soy

iqr_y17_soy = upq_y17_soy - lowq_y17_soy
iqr_y17_soy

miny17_soy = lowq_y17_soy - (0.25*iqr_y17_soy)
miny17_soy

maxy17_soy = upq_y17_soy + (0.25*iqr_y17_soy)
maxy17_soy
```

```{r raw yield quantile cotton 2023}
lowq_y23_cot  = quantile(y23_cotton_w$yield_lbac, 0.1,
                         na.rm = TRUE)
lowq_y23_cot

upq_y23_cot = quantile (y23_cotton_w$yield_lbac, 0.99,
                        na.rm = TRUE)
upq_y23_cot

iqr_y23_cot = upq_y23_cot - lowq_y23_cot
iqr_y23_cot

miny23_cot = lowq_y23_cot - (0.25*iqr_y23_cot)
miny23_cot

maxy23_cot = upq_y23_cot + (0.25*iqr_y23_cot)
maxy23_cot
```

```{r raw yield quantile cotton 2024}
lowq_y24_cot  = quantile(y24_cotton_w$yield_lbac, 0.1,
                         na.rm = TRUE)
lowq_y24_cot

upq_y24_cot = quantile (y24_cotton_w$yield_lbac, 0.99,
                        na.rm = TRUE)
upq_y24_cot

iqr_y24_cot = upq_y24_cot - lowq_y24_cot
iqr_y24_cot

miny24_cot = lowq_y24_cot - (0.25*iqr_y24_cot)
miny24_cot

maxy24_cot = upq_y24_cot + (0.25*iqr_y24_cot)
maxy24_cot
```

## Min/max raw yield quantiles plot

```{r for corn}
ggplot()+
  geom_density(data = y16_corn_w, 
               aes(x = yield_buac)) +
  geom_vline(xintercept = lowq_y16_corn,
             color = "red",
             linetype = 2) +
  geom_vline(xintercept = upq_y16_corn,
             color = "blue",
             linetype = 2) +
  geom_vline(xintercept = miny16_corn,
             color = "red",
             linetype = 1) +
  geom_vline(xintercept = maxy16_corn,
             color = "blue",
             linetype = 1) +
  annotate(geom = "rect",
           ymin = 0,
           ymax = Inf,
           xmin = miny16_corn,
           xmax = maxy16_corn,
           alpha = 0.3,
           fill = "green")


```

## Filtering min/max raw yield

```{r y16_corn y-filtered}
y16_corn_yf = y16_corn_w %>%
  filter(yield_buac > miny16_corn & yield_buac < maxy16_corn)

y16_corn_yf
```

Before filtering we had a total of 23,520 rows. How many rows now after filtering? **22,271** rows

```{r y17_soy y-filtered}
y17_soy_yf = y17_soybeans_w %>%
  filter(yield_buac > miny17_soy & yield_buac < maxy17_soy)

y17_soy_yf
```

Before filtering we had a total of 15,097  rows. How many rows now after filtering? **14,023** rows

```{r y23_cot y-filtered}
y23_cot_yf = y23_cotton_w %>%
  filter(yield_lbac > miny23_cot & yield_lbac < maxy23_cot)

y23_cot_yf
```

Before filtering we had a total of 14,319 rows. How many rows now after filtering? **13,633** rows

```{r y24_cot y-filtered}
y24_cot_yf = y24_cotton_w %>%
  filter(yield_lbac > miny24_cot & yield_lbac < maxy24_cot)

y24_cot_yf
```

Before filtering we had a total of 15,264 rows. How many rows now after filtering? **14,555** rows




## Min/max yield cleaned summary

```{r}
summary(yield20_w_yf$yield_lbac)
```

All right, all done filtering based on yield min and max. Now let’s do the same for speed.

# Yield editor - Min/max speed

## Raw speed distribution

```{r}
ggplot(data = yield20_w, 
       aes(x = speed_mph))+
  geom_density()+
  geom_rug()
```

## Raw speed summary

```{r}
summary(yield20_w$speed_mph) 
```

## Raw speed quantiles

```{r}
lowq_speed = quantile(yield20_w$speed_mph, 0.08)
lowq_speed

upq_speed = quantile(yield20_w$speed_mph, 0.99)
upq_speed

iqr <- upq_speed - lowq_speed
iqr

minv <- lowq_speed - (0.18*iqr)
minv

maxv <- upq_speed + (0.18*iqr)
maxv
```

## Min/max raw speed quantiles plot

```{r}
ggplot()+
  geom_density(data = yield20_w, 
               aes(x = speed_mph))+
  geom_vline(xintercept = lowq_speed, 
             color = "red", 
             linetype = 2)+
  geom_vline(xintercept = upq_speed, 
             color = "blue", 
             linetype = 2)+
  geom_vline(xintercept = minv, 
             color = "red", 
             linetype =1)+
  geom_vline(xintercept = maxv, 
             color = "blue", 
             linetype = 1)+
  annotate(geom = "rect",
           ymin = 0, 
           ymax = Inf,
           xmin = minv, 
           xmax = maxv,
           alpha = .3, 
           fill = "orange")

```

## Filtering min/max raw speed

```{r}
yield20_w_sf = yield20_w_yf %>%
  filter(speed_mph > minv & speed_mph < maxv)

yield20_w_sf
```

## Min/max speed cleaned summary

```{r}
summary(yield20_w_sf$speed_mph)
```

# Yield editor - Position

Original position filter was based on quantiles too. Why filter based on boundary?

## Field boundary

```{r}
boundary = read_sf("../../05-digitizing/ouput/boundary.geojson")

boundary
```

```{r}
boundary_w = boundary %>%
  st_transform(6345)
  
boundary_w
```

## Field boundary plot

```{r}
yieldmap_raw +
  geom_sf(data = boundary_w,
          fill = NA,
          linewidth = 1)
```

Let’s create a negative buffer of 20 m.

## Buffer

```{r}
buffer = boundary_w %>%
  st_buffer(dist = -20) #the unit of whatever is your CRS

buffer
```

```{r}
yieldmap_raw + 
  geom_sf(data=boundary_w, 
          fill = NA,
          linewidth = 1,
          color = "black") +
  geom_sf(data = buffer,
          fill = NA,
          linewidth = 1,
          color = "orange")
```

## Filtering position

```{r}
yield20_w_pf = yield20_w_sf %>%
  st_intersection(buffer)

yield20_w_pf
```

## Position yield cleaned summary

```{r}
summary(yield20_w_pf$yield_lbac)
```

# Raw vs Cleaned data

## Summary statistics

### Raw yield summary

```{r}
summary_raw = yield20_w %>%
  summarise(n = length(yield_lbac),
            min = min(yield_lbac),
            mean = mean(yield_lbac),
            max = max(yield_lbac),
            sd = sd(yield_lbac),
            cv = (sd/mean)*100
            ) %>%
  mutate(data = "raw") %>%
  dplyr::select(data, everything())


summary_raw
```

### Clean yield summary

```{r}
summary_filtered <- yield20_w_pf %>%
  summarise(n = length(yield_lbac),
            min = min(yield_lbac),
            mean = mean(yield_lbac),
            max = max(yield_lbac),
            sd = sd(yield_lbac),
            cv = (sd/mean)*100
            ) %>%
  mutate(data = "clean") %>%
  dplyr::select(data, everything())

summary_filtered

clean_yield20_summary = summary_filtered %>%
  st_drop_geometry() %>%
  select(-data, -n) %>%
  pivot_longer(cols = everything(),
               names_to = "Variable", 
               values_to = "Value") %>%
  mutate(Value = round(Value, 2))

write_csv(clean_yield20_summary, "../ouput/clean_yield20_summary.csv")

read_csv("../ouput/clean_yield20_summary.csv")
```

### Combined yield summary

```{r}
summary_raw %>%
  bind_rows(summary_filtered)
```

## Density plot

```{r}
ggplot()+
  geom_density(data = yield20_w, 
               aes(x = yield_lbac), 
               fill = "blue", 
               alpha = .5) +
  geom_density(data = yield20_w_pf, 
               aes(x = yield_lbac), 
               fill = "yellow", 
               alpha = .5)
```

The density shifted from the bottom left to the peak where mean is.

# Maps

## Raw yield map

```{r}
yieldmap_raw
```

## Clean yield map

```{r}
yieldmap_clean <- ggplot(yield20_w_pf)+
  geom_sf(aes(color = yield_lbac), 
          size = .1)+
  scale_color_viridis_b() +
  labs(color = "Yield (lb/ac)",
       title = "2020 Yield Map") +
  theme_map() +
  theme(legend.position = "right",
        plot.title = element_text(hjust = 0.5)) +
  annotation_scale(location = "bl",
                   width_hint = 0.4,
                   pad_y = unit(-0.0001, "in")) +
  annotation_north_arrow(pad_y = unit(0.1, "in"),
                         style = north_arrow_fancy_orienteering())


yieldmap_clean

ggsave("../ouput/clean_yieldmap20.png",
       bg = "white")
```

## Clean yield map interactive

```{r}
yieldmap_clean_int <- yield20_w_pf %>%
  mapview(zcol = "yield_lbac",
        col.regions = viridis(3),
        alpha.regions = 1,
        cex = .1,
        lwd = 0,
        map.types = "Esri.WorldImagery"
)

yieldmap_clean_int
```

## Side-by-side raw vs. cleaned yield map

```{r}
yieldmap_raw + yieldmap_clean
```

## Side-by-side raw vs. cleaned yield map interactive

```{r}
sync(yieldmap_raw_int, yieldmap_clean_int)
```

# Exporting cleaned version

```{r}
write_sf(yield20_w_pf,
         "../data/yield20_clean.geojson",
         delete_dsn = T)

```

# Summary

Today we:

-   Learned about and implemented 5 yield filters (yield, speed, position)\
-   Compared raw and cleaned yield monitor data

