---
title: "Yield monitor editor - for corn, soybeans and cotton 2016, 2017, 2023, and 2024"
author: "Comfort Adegbenro"
format:
  html:
    #cerulean, journal, quartz, sketchy, slate, morph, superhero, vapor
    self-contained: true
    code-fold: false
    embed-resources: true
    toc: true
    number-sections: true
    theme: sketchy
knitr: 
  opts_chunk: 
    message: false
    warning: false
editor_options: 
  chunk_output_type: inline
---

# Learning objectives

Clean wrangled yield data through different methods, based on IQR of yields (min and max_y), speed (min and max_v), buffer (removing edge effects)
# Setup

## Loading packages

```{r  }
#| message: false

library(readr)
library(janitor)
library(dplyr)
library(tidyr)
library(ggplot2)
library(sf)
library(viridis)
library(mapview)
library(leafsync)
library(patchwork)
library(ggthemes)
library(ggspatial)
```

## Define map theme for re-use
```{r defining map theme}
mytheme_map <- 
  theme_map()+
  theme(legend.position = "bottom",
        legend.direction = "vertical",
        panel.background = element_rect(fill="white",
                                        color=NA),
        plot.background = element_rect(fill="white",
                                       color=NA),
        legend.title = element_text(size=10),
        legend.text = element_text(size=9),
        plot.margin = margin(5, 10, 5, 10))

```


## Importing data

```{r import all yield data for all crops and boundary}
y16_corn_w = read_sf("../data/y16corn_processed.geojson")
y16_corn_w

y17_soybeans_w = read_sf("../data/y17soybeans_processed.geojson")
y17_soybeans_w

y23_cotton_w = read_sf("../data/y23cotton_processed.geojson")
y23_cotton_w

y24_cotton_w = read_sf("../data/y24cotton_processed.geojson")
y24_cotton_w

boundary = read_sf("../data/50ac/boundary/BackFieldPivot.shp") %>%
  st_transform(crs = 6345)

boundary
```

# EDA

## Summary

```{r summaries of yields for all crops }
summary(y16_corn_w)
summary(y17_soybeans_w)
summary(y23_cotton_w)
summary(y24_cotton_w)
```

## Raw map

```{r raw maps for corn}
ymap16_corn_raw =  y16_corn_w %>%
  filter(yield_buac < 300) %>%
  ggplot()+
  geom_sf(aes(color = yield_buac),
          size = 0.3) +
  scale_color_viridis_b() +
  #geom_sf(data = boundary, fill = NA) +
  labs(color = "Corn yield (bu/acre)",
       title = "Map of Corn Raw Yield Data (outlier of >300 bu/ac excluded)") +
  mytheme_map +
  annotation_scale(location = "bl",
                   width_hint = 0.3,
                   pad_x = unit(0.1, "in"),
                   pad_y = unit(0.01, "in")) +
  annotation_north_arrow(location = "tl",
                       pad_x = unit(0.1, "in"),
                       pad_y = unit(0.05, "in"),
                       height = unit(0.4, "in"),
                       width = unit(0.4, "in"),
                       style = north_arrow_fancy_orienteering())
ymap16_corn_raw
```

# Yield editor - Min/max yield

## Raw yield quantiles

```{r raw yield quantile corn 2016}
lowq_y16_corn  = quantile(y16_corn_w$yield_buac, 0.1)
lowq_y16_corn

upq_y16_corn = quantile (y16_corn_w$yield_buac, 0.99)
upq_y16_corn

iqr_y16_corn = upq_y16_corn - lowq_y16_corn
iqr_y16_corn

miny16_corn = lowq_y16_corn - (0.25*iqr_y16_corn)
miny16_corn

maxy16_corn = upq_y16_corn + (0.25*iqr_y16_corn)
maxy16_corn
```


```{r raw yield quantile soybean 2017}
lowq_y17_soy  = quantile(y17_soybeans_w$yield_buac, 0.1)
lowq_y17_soy

upq_y17_soy = quantile (y17_soybeans_w$yield_buac, 0.99)
upq_y17_soy

iqr_y17_soy = upq_y17_soy - lowq_y17_soy
iqr_y17_soy

miny17_soy = lowq_y17_soy - (0.25*iqr_y17_soy)
miny17_soy

maxy17_soy = upq_y17_soy + (0.25*iqr_y17_soy)
maxy17_soy
```

```{r raw yield quantile cotton 2023}
lowq_y23_cot  = quantile(y23_cotton_w$yield_lbac, 0.1,
                         na.rm = TRUE)
lowq_y23_cot

upq_y23_cot = quantile (y23_cotton_w$yield_lbac, 0.99,
                        na.rm = TRUE)
upq_y23_cot

iqr_y23_cot = upq_y23_cot - lowq_y23_cot
iqr_y23_cot

miny23_cot = lowq_y23_cot - (0.25*iqr_y23_cot)
miny23_cot

maxy23_cot = upq_y23_cot + (0.25*iqr_y23_cot)
maxy23_cot
```

```{r raw yield quantile cotton 2024}
lowq_y24_cot  = quantile(y24_cotton_w$yield_lbac, 0.1,
                         na.rm = TRUE)
lowq_y24_cot

upq_y24_cot = quantile (y24_cotton_w$yield_lbac, 0.99,
                        na.rm = TRUE)
upq_y24_cot

iqr_y24_cot = upq_y24_cot - lowq_y24_cot
iqr_y24_cot

miny24_cot = lowq_y24_cot - (0.25*iqr_y24_cot)
miny24_cot

maxy24_cot = upq_y24_cot + (0.25*iqr_y24_cot)
maxy24_cot
```

## Min/max raw yield quantiles plot

```{r for corn}
ggplot()+
  geom_density(data = y16_corn_w, 
               aes(x = yield_buac)) +
  geom_vline(xintercept = lowq_y16_corn,
             color = "red",
             linetype = 2) +
  geom_vline(xintercept = upq_y16_corn,
             color = "blue",
             linetype = 2) +
  geom_vline(xintercept = miny16_corn,
             color = "red",
             linetype = 1) +
  geom_vline(xintercept = maxy16_corn,
             color = "blue",
             linetype = 1) +
  annotate(geom = "rect",
           ymin = 0,
           ymax = Inf,
           xmin = miny16_corn,
           xmax = maxy16_corn,
           alpha = 0.3,
           fill = "green")


```

## Filtering min/max raw yield

```{r y16_corn y-filtered}
y16_corn_yf = y16_corn_w %>%
  filter(yield_buac > miny16_corn & yield_buac < maxy16_corn)

y16_corn_yf
```

Before filtering we had a total of 23,520 rows. How many rows now after filtering? **22,271** rows

```{r y17_soy y-filtered}
y17_soy_yf = y17_soybeans_w %>%
  filter(yield_buac > miny17_soy & yield_buac < maxy17_soy)

y17_soy_yf
```

Before filtering we had a total of 15,097  rows. How many rows now after filtering? **14,023** rows

```{r y23_cot y-filtered}
y23_cot_yf = y23_cotton_w %>%
  filter(yield_lbac > miny23_cot & yield_lbac < maxy23_cot)

y23_cot_yf
```

Before filtering we had a total of 14,319 rows. How many rows now after filtering? **13,633** rows

```{r y24_cot y-filtered}
y24_cot_yf = y24_cotton_w %>%
  filter(yield_lbac > miny24_cot & yield_lbac < maxy24_cot)

y24_cot_yf
```

Before filtering we had a total of 15,264 rows. How many rows now after filtering? **14,555** rows


## Min/max yield cleaned summary

```{r using cotton 2023 as example}
summary(y23_cot_yf$yield_lbac)
```

Now filter based on speed.

# Yield editor - Min/max speed

## Raw speed quantiles

```{r raw speed quantile corn 2016}
lowq_s16_corn  = quantile(y16_corn_w$speed_mph, 0.08)
lowq_s16_corn

upq_s16_corn = quantile (y16_corn_w$speed_mph, 0.99)
upq_s16_corn

iqr_s16_corn = upq_s16_corn - lowq_s16_corn
iqr_s16_corn

minv16_corn = lowq_s16_corn - (0.18*iqr_s16_corn)
minv16_corn

maxv16_corn = upq_s16_corn + (0.18*iqr_s16_corn)
maxv16_corn
```


```{r raw speed quantile soybean 2017}
lowq_s17_soy  = quantile(y17_soybeans_w$speed_mph, 0.08)
lowq_s17_soy

upq_s17_soy = quantile (y17_soybeans_w$speed_mph, 0.99)
upq_s17_soy

iqr_s17_soy = upq_s17_soy - lowq_s17_soy
iqr_s17_soy

minv17_soy = lowq_s17_soy - (0.18*iqr_s17_soy)
minv17_soy

maxv17_soy = upq_s17_soy + (0.18*iqr_s17_soy)
maxv17_soy
```

```{r raw speed quantile cotton 2023}
lowq_s23_cot  = quantile(y23_cotton_w$speed_mph, 0.08,
                         na.rm = TRUE)
lowq_s23_cot

upq_s23_cot = quantile (y23_cotton_w$speed_mph, 0.99,
                        na.rm = TRUE)
upq_s23_cot

iqr_s23_cot = upq_s23_cot - lowq_s23_cot
iqr_s23_cot

minv23_cot = lowq_s23_cot - (0.18*iqr_s23_cot)
minv23_cot

maxv23_cot = upq_s23_cot + (0.18*iqr_s23_cot)
maxv23_cot
```


```{r raw speed quantile cotton 2024}
lowq_s24_cot  = quantile(y24_cotton_w$speed_mph, 0.08,
                         na.rm = TRUE)
lowq_s24_cot

upq_s24_cot = quantile (y24_cotton_w$speed_mph, 0.99,
                        na.rm = TRUE)
upq_s24_cot

iqr_s24_cot = upq_s24_cot - lowq_s24_cot
iqr_s24_cot

minv24_cot = lowq_s24_cot - (0.18*iqr_s24_cot)
minv24_cot

maxv24_cot = upq_s24_cot + (0.18*iqr_s24_cot)
maxv24_cot
```


## Min/max raw speed quantiles plot

```{r using cotton 2023 as example}
ggplot()+
  geom_density(data = y23_cotton_w, 
               aes(x = speed_mph))+
  geom_vline(xintercept = lowq_s23_cot, 
             color = "red", 
             linetype = 2)+
  geom_vline(xintercept = upq_s23_cot, 
             color = "blue", 
             linetype = 2)+
  geom_vline(xintercept = minv23_cot, 
             color = "red", 
             linetype =1)+
  geom_vline(xintercept = maxv23_cot, 
             color = "blue", 
             linetype = 1)+
  annotate(geom = "rect",
           ymin = 0, 
           ymax = Inf,
           xmin = minv23_cot, 
           xmax = maxv23_cot,
           alpha = .3, 
           fill = "orange")

```

## Filtering min/max raw speed

```{r}
yield20_w_sf = yield20_w_yf %>%
  filter(speed_mph > minv & speed_mph < maxv)

yield20_w_sf
```


```{r y16_corn s-filtered}
y16_corn_sf = y16_corn_yf %>%
  filter(speed_mph > minv16_corn & speed_mph < maxv16_corn)

y16_corn_sf
```

Before filtering we had a total of 22,271 rows. How many rows now after filtering? **22,003** rows

```{r y17_soy s-filtered}
y17_soy_sf = y17_soy_yf %>%
  filter(speed_mph > minv17_soy & speed_mph < maxv17_soy)

y17_soy_sf
```

Before filtering we had a total of 14,023  rows. How many rows now after filtering? **13,673** rows

```{r y23_cot s-filtered}
y23_cot_sf = y23_cot_yf %>%
  filter(speed_mph > minv23_cot & speed_mph < maxv23_cot)

y23_cot_sf
```

Before filtering we had a total of 13,633 rows. How many rows now after filtering? **13,220** rows

```{r y24_cot s-filtered}
y24_cot_sf = y24_cot_yf %>%
  filter(speed_mph > minv24_cot & speed_mph < maxv24_cot)

y24_cot_sf
```

Before filtering we had a total of 14,555 rows. How many rows now after filtering? **14,414** rows

## Min/max speed cleaned summary

```{r cotton 23 as example}
summary(y23_cot_sf$speed_mph)
```

# Yield editor - Position

## Field boundary plot

```{r ymap corn 16 with boundary}
ymap16_corn_raw +
  geom_sf(data = boundary,
          fill = NA,
          linewidth = 1)
```

Creating a negative buffer of 20 m.

## Buffer

```{r}
buffer = boundary %>%
  st_buffer(dist = -20) #the unit of CRS

buffer
```

```{r}
ymap16_corn_raw + 
  geom_sf(data=boundary, 
          fill = NA,
          linewidth = 1,
          color = "black") +
  geom_sf(data = buffer,
          fill = NA,
          linewidth = 1,
          color = "orange")
```

## Filtering position

```{r y16_corn buffer filtered}
y16_corn_bf = y16_corn_sf %>%
  st_intersection(buffer)

y16_corn_bf
```


```{r y17_soy buffer filtered}
y17_soy_bf = y17_soy_sf %>%
  st_intersection(buffer)

y17_soy_bf
```


```{r y23_cot buffer filtered}
y23_cot_bf = y23_cot_sf %>%
  st_intersection(buffer)

y23_cot_bf
```


```{r y24_cot buffer filtered}
y24_cot_bf = y24_cot_sf %>%
  st_intersection(buffer)

y24_cot_bf
```

## Position yield cleaned summary

```{r}
summary(y23_cot_bf$yield_lbac)
```

# Cleaned data

## Summary statistics

### Clean yield summary

# Creating Summaries
```{r summary data y16 corn}
summary_raw_y16_corn = y16_corn_all_w %>%
  st_drop_geometry() %>%
  summarise(Observations = length(yield_buac),
            Min = min(yield_buac),
            Mean = mean(yield_buac),
            Max = max(yield_buac),
            Std_Dev = sd(yield_buac)
            ) %>%
  mutate(Data = "Raw",
         Crop = "Corn",
         Year = 2016) %>%
  dplyr::select(Data, Year, Crop, everything())

summary_raw_y16_corn
```

```{r summary data y17 soybeans}
summary_raw_y17_soybeans = y17_soybeans_w %>%
  st_drop_geometry() %>%
  summarise(Observations = length(yield_buac),
            Min = min(yield_buac),
            Mean = mean(yield_buac),
            Max = max(yield_buac),
            Std_Dev = sd(yield_buac)
            ) %>%
  mutate(Data = "Raw",
         Crop = "Soybeans",
         Year = 2017) %>%
  dplyr::select(Data, Year, Crop, everything())

summary_raw_y17_soybeans
```

```{r summary data y23 cotton}
summary_raw_y23_cotton = y23_cotton_w %>%
  st_drop_geometry() %>%
  summarise(Observations = length(yield_lbac),
            Min = min(yield_lbac, na.rm = TRUE),
            Mean = mean(yield_lbac, na.rm = TRUE),
            Max = max(yield_lbac, na.rm = TRUE),
            Std_Dev = sd(yield_lbac, na.rm = TRUE)
            ) %>%
  mutate(Data = "Raw",
         Crop = "Cotton",
         Year = 2023) %>%
  dplyr::select(Data, Year, Crop, everything())

summary_raw_y23_cotton
```

```{r summary data y24 cotton}
summary_raw_y24_cotton = y24_cotton_w %>%
  st_drop_geometry() %>%
  summarise(Observations = length(yield_lbac),
            Min = min(yield_lbac, na.rm = TRUE),
            Mean = mean(yield_lbac, na.rm = TRUE),
            Max = max(yield_lbac, na.rm = TRUE),
            Std_Dev = sd(yield_lbac, na.rm = TRUE)
            ) %>%
  mutate(Data = "Raw",
         Crop = "Cotton",
         Year = 2024) %>%
  dplyr::select(Data, Year, Crop, everything())

summary_raw_y24_cotton
```


```{r combine all yield summaries}
all_ys = summary_raw_y16_corn %>%
  bind_rows(summary_raw_y17_soybeans) %>%
  bind_rows(summary_raw_y23_cotton) %>%
  bind_rows(summary_raw_y24_cotton)
  
all_ys

all_ys_table = all_ys %>%
  gt() %>%
  tab_header(title = "Yield Summary: Corn, Soybeans and Cotton") %>%
  fmt_number(columns = c(Mean, Std_Dev, Min, Max), decimals = 1) %>%
  fmt_number(columns = Observations, decimals = 0, use_seps = TRUE) %>%
  tab_source_note(source_note = "Yield units: bu/ac for corn and soybean, and lb/ac for cotton") %>%
  tab_options(
    table.width = pct(90),
    table.font.size = 14,
    data_row.padding = px(8),
    column_labels.font.weight = "bold",
    heading.title.font.size = 18,
    heading.title.font.weight = "bold"
  )

all_ys_table

# Saving to load it in another script
saveRDS(all_ys_table, "../output/all_ys_table.rds")
```



## Density plot

```{r}
ggplot()+
  geom_density(data = yield20_w, 
               aes(x = yield_lbac), 
               fill = "blue", 
               alpha = .5) +
  geom_density(data = yield20_w_pf, 
               aes(x = yield_lbac), 
               fill = "yellow", 
               alpha = .5)
```

The density shifted from the bottom left to the peak where mean is.

# Maps

## Raw yield map

```{r}
yieldmap_raw
```

## Clean yield map

```{r}
yieldmap_clean <- ggplot(yield20_w_pf)+
  geom_sf(aes(color = yield_lbac), 
          size = .1)+
  scale_color_viridis_b() +
  labs(color = "Yield (lb/ac)",
       title = "2020 Yield Map") +
  theme_map() +
  theme(legend.position = "right",
        plot.title = element_text(hjust = 0.5)) +
  annotation_scale(location = "bl",
                   width_hint = 0.4,
                   pad_y = unit(-0.0001, "in")) +
  annotation_north_arrow(pad_y = unit(0.1, "in"),
                         style = north_arrow_fancy_orienteering())


yieldmap_clean

ggsave("../ouput/clean_yieldmap20.png",
       bg = "white")
```


# Exporting cleaned version

```{r}
write_sf(yield20_w_pf,
         "../data/yield20_clean.geojson",
         delete_dsn = T)

```

