---
author: "Comfort Adegbenro"
title: "Clean Yield Interpolation for Corn, Soybeans and Cotton"
format: html
---

# Objectives  

Interpolate clean yield data set using Inverse distance weighted (IDW), and then extract the interpolated yield data, and save file.



# Setup  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#install.packages("gstat")
#install.packages("stars")
#install.packages("cubelyr")

# Packages
library(dplyr)
library(tidyr)
library(sf)
library(ggplot2)
library(viridis)
library(gstat)
library(raster)
library(stars)
library(tidyverse)
library(ggthemes)
library(ggspatial)
library(patchwork)
```


```{r defining map theme}
mytheme_map <- 
  theme_map()+
  theme(legend.position = "bottom",
        legend.direction = "vertical",
        panel.background = element_rect(fill="white",
                                        color=NA),
        plot.background = element_rect(fill="white",
                                       color=NA),
        legend.title = element_text(size=10),
        legend.text = element_text(size=9),
        plot.margin = margin(5, 10, 5, 10))

```


```{r reading yield data}
y16_corn_c = read_sf("../data/y16_corn_clean.geojson")
y16_corn_c

y17_soy_c = read_sf("../data/y17_soy_clean.geojson")
y17_soy_c

y23_cot_c = read_sf("../data/y23_cot_clean.geojson")
y23_cot_c

y24_cot_c = read_sf("../data/y24_cot_clean.geojson")
y24_cot_c


summary(y16_corn_c$elev_ft)
summary(y17_soy_c$elev_ft)
summary(y23_cot_c$elev_ft)
summary(y24_cot_c$elev_ft)
```


```{r boundary, echo=F}
boundary = read_sf("../data/50ac/boundary/BackFieldPivot.shp") %>%
  st_transform(crs = 6345)

boundary
```

# Interpolation  

## Grid
```{r grid_v}
# grid in vector format
grid_r = st_make_grid(boundary,
                      cellsize = 5) %>%
  st_as_sf() %>%
  st_rasterize(dx = 5, dy = 5) %>%
  st_crop(boundary)

grid_r
```
## IDW  
Let's set up the IDW algorithm by providing a formula. Here, we are just setting up the model.    
```{r idw_mod}
y16corn_idw_mod = gstat(data = y16_corn_c,
                formula = yield_buac ~ 1)

y17soy_idw_mod = gstat(data = y17_soy_c,
                formula = yield_buac ~ 1)

y23soy_idw_mod = gstat(data = y23_cot_c,
                formula = yield_lbac ~ 1)

y24soy_idw_mod = gstat(data = y24_cot_c,
                formula = yield_lbac ~ 1)

```

Now, we need to actually predict our IDW model on our raster grid (this is where the interpolation really happens).  
```{r idw_pred}
y16corn_idw_pred = predict(y16corn_idw_mod, grid_r)
y16corn_idw_pred

y17soy_idw_pred = predict(y17soy_idw_mod, grid_r)
y17soy_idw_pred

y23cot_idw_pred = predict(y23soy_idw_mod, grid_r)
y23cot_idw_pred

y24cot_idw_pred = predict(y24soy_idw_mod, grid_r)
y24cot_idw_pred

class(y24cot_idw_pred)
```


# Grid extraction  
In this step, the goal is to extract the interpolated yield value from the raster and basically transform it from cell (raster) to a polygon (vector).  

```{r idw_pred_v}
y16corn_idw_pred_v = y16corn_idw_pred %>%
  st_as_sf(as_points = F,
           merge = F) %>%
  dplyr::select(corn_ipyield_buac = var1.pred)
y16corn_idw_pred_v


y17soy_idw_pred_v = y17soy_idw_pred %>%
  st_as_sf(as_points = F,
           merge = F) %>%
  dplyr::select(soy_ipyield_buac = var1.pred)
y17soy_idw_pred_v

y23cot_idw_pred_v = y23cot_idw_pred %>%
  st_as_sf(as_points = F,
           merge = F) %>%
  dplyr::select(cot23_ipyield_lbac = var1.pred)
y23cot_idw_pred_v

y24cot_idw_pred_v = y24cot_idw_pred %>%
  st_as_sf(as_points = F,
           merge = F) %>%
  dplyr::select(cot24_ipyield_lbac = var1.pred)
y24cot_idw_pred_v

class(y16corn_idw_pred_v)
```


# Interpolated yield map

```{r ip corn map}
ymap16_corn_ip <- ggplot()+
  geom_sf(data = y16corn_idw_pred_v,
          aes(fill = corn_ipyield_buac),
          color = NA,
          size = .1)+
  scale_fill_viridis_b() +
  geom_sf(data = boundary, fill = NA,
          linewidth = 1) +
  labs(fill = "Yield (bu/ac)",
       title = "2016 Corn") +
  mytheme_map +
  annotation_scale(location = "bl",
                   width_hint = 0.3,
                   pad_x = unit(0.1, "in"),
                   pad_y = unit(0.01, "in")) +
  annotation_north_arrow(location = "tl",
                       pad_x = unit(0.1, "in"),
                       pad_y = unit(0.05, "in"),
                       height = unit(0.4, "in"),
                       width = unit(0.4, "in"),
                       style = north_arrow_fancy_orienteering())


ymap16_corn_ip
```


```{r ip soybean map}
ymap17_soy_ip <- ggplot()+
  geom_sf(data = y17soy_idw_pred_v,
          aes(fill = soy_ipyield_buac), 
          color = NA,
          size = .1)+
  scale_fill_viridis_b() +
  geom_sf(data = boundary, fill = NA,
          linewidth = 1) +
  labs(fill = "Yield (bu/ac)",
       title = "2017 Soybeans") +
  mytheme_map


ymap17_soy_ip
```


```{r ip cotton 23 map}
ymap23_cot_ip <- ggplot()+
  geom_sf(data = y23cot_idw_pred_v,
          aes(fill = cot23_ipyield_lbac),
          color = NA,
          size = .1)+
  scale_fill_viridis_b() +
  geom_sf(data = boundary, fill = NA,
          linewidth = 1) +
  labs(fill = "Yield (lb/ac)",
       title = "2023 Cotton") +
  mytheme_map


ymap23_cot_ip
```


```{r ip cotton 24 map}
ymap24_cot_ip <- ggplot()+
  geom_sf(data = y24cot_idw_pred_v,
          aes(fill = cot24_ipyield_lbac), 
          color = NA,
          size = .1)+
  scale_fill_viridis_b() +
  geom_sf(data = boundary, fill = NA,
          linewidth = 1) +
  labs(fill = "Yield (lb/ac)",
       title = "2024 Cotton") +
  mytheme_map


ymap24_cot_ip
```

```{r combine ip yield maps}
ymap16_corn_ip + ymap17_soy_ip + ymap23_cot_ip + ymap24_cot_ip +
  plot_layout(ncol = 4) +
  plot_annotation(title = "Yield Maps of Interpolated Data",
    theme = theme(
      plot.title = element_text(size = 13, face = "bold", hjust = 0.5)))

ggsave("../output/ip-yield_maps.png",
       width = 18,
       height = 5)

```


# Exporting interpolated version

```{r export all clean yield data}
write_sf(y16corn_idw_pred_v, "../data/y16_corn_ip.geojson", delete_dsn = T)
write_sf(y17soy_idw_pred_v, "../data/y17_soy_ip.geojson", delete_dsn = T)
write_sf(y23cot_idw_pred_v, "../data/y23_cot_ip.geojson", delete_dsn = T)
write_sf(y24cot_idw_pred_v, "../data/y24_cot_ip.geojson", delete_dsn = T)
```