---
title: "Interpolation yield20"
format: html
---

# Learning Objectives  
Today, our learning objectives are:  
- To learn about why is interpolation needed and useful  
- Explore one type of interpolation analysis (inverse distance weighted - IDW)  
- Perform IDW on the yield data sets  
- Extract interpolated yield data based on a grid and save to file  


# Setup  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#install.packages("gstat")
#install.packages("stars")
#install.packages("cubelyr")

# Packages
library(dplyr)
library(tidyr)
library(sf)
library(ggplot2)
library(viridis)
library(gstat)
library(raster)
library(stars)
library(tidyverse)
library(ggthemes)
library(ggspatial)
```

```{r reading yield data}
yield20_c = read_sf("../data/yield20_clean.geojson")

yield20_c
```

```{r summary}
summary(yield20_c)
```


```{r boundary, echo=F}
boundary_w = read_sf("../../05-digitizing/ouput/boundary.geojson") %>%
  st_transform(crs = 6345)

boundary_w
```

# Interpolation  
Why do we need to interpolate?  

```{r yield ip map}
yieldip_map <- ggplot()+
  geom_sf(data = yield20_c,
          aes(color=yield_lbac), size=.1)+
  scale_color_viridis()+
  geom_sf(data = boundary_w,
          fill=NA,
          color="blue")

yieldip_map
```

Interpolation methods:  
  - deterministic: ex. **inverse distance weight (IDW)**  
  - statistical: ex. kriging  

For simplicity sake, we are going to use IDW for this exercise.  

## Grid
```{r grid_v}
# grid in vector format
grid_v = st_make_grid(boundary_w,
                      cellsize = 10)

grid_v

class(grid_v)
```

```{r grid vector plot}
yieldip_map +
  geom_sf(data = grid_v,
          alpha = 0.5,
          fill = NA,
          size = 0.1)

```

One issue here is that interpolation methods in R do not accept grids in vector format, only in **raster** format.  

Let's rasterize our vector grid.  

```{r grid_r}
# grid in raster format 
grid_r = grid_v %>%
  st_as_sf() %>%
  st_rasterize(dx = 10, dy = 10) %>%
  st_crop(boundary_w)

grid_r

class(grid_r)
```

```{r grid_r plot}
yieldip_map +
  geom_stars(data = grid_r,
             alpha = .6) 
```

## IDW  
Let's set up the IDW algorithm by providing a formula. Here, we are just setting up the model.    
```{r idw_mod}
idw_mod = gstat(data = yield20_c,
                formula = yield_lbac ~ 1)

idw_mod

class(idw_mod)
```

Now, we need to actually predict our IDW model on our raster grid (this is where the interpolation really happens).  
```{r idw_pred}
idw_pred = predict(idw_mod, grid_r)

idw_pred

class(idw_pred)
```

```{r idw_pred plot}
ggplot() +
  geom_stars(data = idw_pred) +
  scale_fill_viridis() +
  geom_sf(data = boundary_w, 
          fill = NA) +
  geom_sf(data = grid_v,
          alpha = 0.5,
          fill = NA,
          size = 0.1)

```

# Grid extraction  
In this step, the goal is to extract the interpolated yield value from the raster and basically transform it from cell (raster) to a polygon (vector).  

```{r idw_pred_v}
idw_pred_v = idw_pred %>%
  st_as_sf(as_points = F,
           merge = F) %>%
  dplyr::select(ipyield_lbac = var1.pred)

idw_pred_v

class(idw_pred_v)
```

```{r idw_pred_v plot}
ggplot() +
  geom_sf(data = idw_pred_v,
          aes(fill = ipyield_lbac),
          color = NA) +
  scale_fill_viridis()
```

```{r interpolated distribution}
ggplot(data = idw_pred_v) +
  geom_density(aes(x = ipyield_lbac)) +
  geom_density(data = yield20_c,
               aes(x=yield_lbac),
               color = "blue")

```

# Exporting to file  

Since we will use this product in a future class (to look into yield stability and create management zones), let's export it to file.  

```{r exporting}
write_sf(idw_pred_v,
         "../data/yield20_interpolated.geojson",
         delete_dsn = T)
```

# Assignment #6  

```{r}
summary_filtered <- idw_pred_v %>%
  summarise(min = min(ipyield_lbac),
            median = median(ipyield_lbac),
            max = max(ipyield_lbac),
            ) %>%
  st_drop_geometry() %>%
  pivot_longer(cols = everything(),
               names_to = "Variable", 
               values_to = "Value") %>%
  mutate(Value = round(Value, 2))

write_csv(summary_filtered, "../ouput/ipyield20_summary.csv")

```


```{r}
yieldmap_ip <- ggplot(idw_pred_v)+
  geom_sf(aes(color = ipyield_lbac), 
          size = 1.3)+
  scale_color_viridis_b() +
  labs(color = "Yield (lb/ac)",
       title = "2020 Interpolated Yield Map") +
  theme_map() +
  theme(legend.position = "right",
        plot.title = element_text(hjust = 0.5)) +
  annotation_scale(location = "bl",
                   width_hint = 0.4,
                   pad_y = unit(-0.0001, "in")) +
  annotation_north_arrow(pad_y = unit(0.1, "in"),
                         style = north_arrow_fancy_orienteering())


yieldmap_ip

ggsave("../ouput/ip_yieldmap20.png",
       bg = "white")
```

