---
title: "Yield stability"
author: "Comfort Adegbenro"
format:
  html:
    code-fold: false
    embed-resources: true
    toc: true
    number-sections: true
    theme: journal
    #cerulean, journal, quartz, sketchy, slate, morph, superhero, vapor
knitr: 
  opts_chunk: 
    message: false
    warning: false
---

# Objectives  
The goal is to import the clean and interpolated yield maps for the three crops and 4 years, create year/crop specific spatial yield stability areas, then create spatio-temporal yield stability areas, and export the yield stability layer.  


# Setup  
```{r setup }
# Packages
library(dplyr)
library(tidyr)
library(sf)
library(ggplot2)
library(viridis)
library(ggthemes)
library(tibble)
library(patchwork)
library(gt)
```


```{r defining map theme}
mytheme_map <- 
  theme_map()+
  theme(legend.position = "bottom",
        legend.direction = "vertical",
        panel.background = element_rect(fill="white",
                                        color=NA),
        plot.background = element_rect(fill="white",
                                       color=NA),
        legend.title = element_text(size=10),
        legend.text = element_text(size=9),
        plot.margin = margin(5, 10, 5, 10))

```

## Data import
```{r importing all interpolated yields}
y16corn_ip = read_sf("../data/y16_corn_ip.geojson")

y17soy_ip = read_sf("../data/y17_soy_ip.geojson")

y23cot_ip = read_sf("../data/y23_cot_ip.geojson")

y24cot_ip = read_sf("../data/y24_cot_ip.geojson")

boundary = read_sf("../data/50ac/boundary/BackFieldPivot.shp") %>%
  st_transform(crs = 6345)

summary(y16corn_ip)
summary(y17soy_ip)
summary(y23cot_ip)
summary(y24cot_ip)
```


# Spatial stability  
We have yield data for the same field, but different crops, and different years, so we standardize by median for each year (/crop) which is more stable compared to mean. Calculate the median and then divide the yield observations for each year by that year's median.


Beginning with 2017 yield data.
```{r sy_16c}
sy_16c = y16corn_ip %>%
  mutate(medianyield_buac = median(corn_ipyield_buac)) %>%
  mutate(sy = corn_ipyield_buac/medianyield_buac *100) %>%
  mutate(sy = round(sy, 1)) %>%
  dplyr::select(sy16c = sy, geometry)

sy_16c
summary(sy_16c)
```
Let's check how does this standardization looks like in space.  

```{r sy_16c map}
sy_16c_map = ggplot() +
  geom_sf(data = sy_16c, 
          aes(fill = sy16c),
          color = NA) +
  scale_fill_viridis_b(option = "D", 
                       breaks = seq(0, 160, 20),
                       limits = c(0, 160)) +
   geom_sf(data = boundary, fill = NA,
          linewidth = 1) +
 # scale_fill_gradient2(low = "#440154FF", 
                #       mid = "white",
               #        high = "#22A884FF",
              #         midpoint = 100,
               #        breaks = seq(0, 160, 20),
               #        limits = c(0, 160)) +
  labs(title = "Corn (2016)",
       fill = "Median-\nstandardized\nyield") +
  mytheme_map +
  annotation_scale(location = "bl",
                   width_hint = 0.3,
                   pad_x = unit(0.1, "in"),
                   pad_y = unit(0.01, "in")) +
  annotation_north_arrow(location = "tl",
                       pad_x = unit(0.1, "in"),
                       pad_y = unit(0.05, "in"),
                       height = unit(0.4, "in"),
                       width = unit(0.4, "in"),
                       style = north_arrow_fancy_orienteering())

sy_16c_map 
```
Repeat the same steps for other years.  

```{r sy_17s}
sy_17s = y17soy_ip %>%
  mutate(medianyield_buac = median(soy_ipyield_buac)) %>%
  mutate(sy = soy_ipyield_buac/medianyield_buac *100) %>%
  mutate(sy = round(sy, 1)) %>%
  dplyr::select(sy17s = sy, geometry)


sy_17s

summary(sy_17s)
```

```{r sy_17s map}
sy_17s_map <- ggplot()+
  geom_sf(data = sy_17s,
          aes(fill = sy17s), 
          color = NA)+
  scale_fill_viridis_b(option = "D", 
                       breaks = seq(0, 160, 20),
                       limits = c(0, 160)) +
   geom_sf(data = boundary, fill = NA,
          linewidth = 1) +
  labs(title="Soybeans (2017)")+
  mytheme_map +
  theme(legend.position = "none")


sy_17s_map 
```
  

```{r sy_23c}
sy_23c  = y23cot_ip %>%
  mutate(medianyield_lbac = median(cot23_ipyield_lbac)) %>%
  mutate(sy = cot23_ipyield_lbac/medianyield_lbac *100) %>%
  mutate(sy = round(sy, 1)) %>%
  dplyr::select(sy23c = sy, geometry)


sy_23c
summary(sy_23c)
```

```{r sy_23c map}
sy_23c_map <- ggplot()+
  geom_sf(data = sy_23c,
          aes(fill = sy23c), 
          color = NA)+
   scale_fill_viridis_b(option = "D", 
                       breaks = seq(0, 160, 20),
                       limits = c(0, 160)) +
   geom_sf(data = boundary, fill = NA,
          linewidth = 1) +
  labs(title="Cotton (2023)") +
  mytheme_map +
  theme(legend.position = "none")

sy_23c_map 
```

```{r sy_24c}
sy_24c  = y24cot_ip %>%
  mutate(medianyield_lbac = median(cot24_ipyield_lbac)) %>%
  mutate(sy = cot24_ipyield_lbac/medianyield_lbac *100) %>%
  mutate(sy = round(sy, 1)) %>%
  dplyr::select(sy24c = sy, geometry)


sy_24c
summary(sy_24c)
```


```{r sy_24c map}
sy_24c_map <- ggplot()+
  geom_sf(data = sy_24c,
          aes(fill = sy24c), 
          color = NA)+
   scale_fill_viridis_b(option = "D", 
                       breaks = seq(0, 160, 20),
                       limits = c(0, 160)) +
   geom_sf(data = boundary, fill = NA,
          linewidth = 1) +
  labs(title="Cotton (2024)") +
  mytheme_map +
  theme(legend.position = "none")

sy_24c_map 
```


Plot all 4 maps side-by-side to assist in visualizing them.  

```{r all maps}
library(patchwork)
sy_16c_map + sy_17s_map + sy_23c_map + sy_24c_map +
  plot_layout(ncol = 4) +
  plot_annotation(title = "Yield Maps of Spatial Stability",
    theme = theme(
      plot.title = element_text(size = 13, face = "bold", hjust = 0.5)))

ggsave("../output/sp_stab-yield_maps.png",
       width = 18,
       height = 5)

```


```{r summary after spatial stability}
summary(sy_16c$sy16c) #60 to 152
summary(sy_17s$sy17s) #60 to 119
summary(sy_23c$sy23c) #68 to 143
summary(sy_24c$sy24c) #54 to 140
```

# Spatio-temporal stability  
To perform a temporal analysis, we need to merge all three data frames into one.  

After joining the data sets, we still only have their information on year-specific yield spatial standardization. 

> How can we merge these pieces of information to create a spatial-temporal classification?  

We can use the yield mean and coefficient of variation (CV) of each pixel over 4 years, and use that to guide our decision.    

First, let's merge the 4 data sets and calculate pixel-based mean and cv.

## Pixel-based mean and cv  
```{r}
sy_all = sy_16c %>%
  st_join(sy_17s,
          join = st_equals,
          left = T)%>%
  st_join(sy_23c,
          join = st_equals,
          left = T) %>%
  st_join(sy_24c,
          join = st_equals,
          left = T) %>%
  rowwise() %>%
  mutate(mean_pixel = mean(c(sy16c, sy17s, sy23c, sy24c))) %>%
  mutate(sd_pixel = sd(c(sy16c, sy17s, sy23c, sy24c))) %>%
  mutate(cv_pixel = sd_pixel/mean_pixel*100)

sy_all

```

## EDA on pixel-based stats  
Summary:  
```{r summary }
summary(sy_all$cv_pixel)
```

Density plot:
```{r density plot }
ggplot(data = sy_all,
       aes(x = mean_pixel)) +
  geom_density() +
  geom_rug()

```


```{r find quantiles to define mean and cv marks}
quantile(sy_all$mean_pixel, 0.15, na.rm = TRUE)
quantile(sy_all$mean_pixel, 0.95, na.rm = TRUE)
```


Standardized mean map
```{r mean map}
sy_all_mean = ggplot() +
  geom_sf(data = sy_all,
          aes(fill = mean_pixel),
          color = NA) +
  scale_fill_viridis_b(option = "C",
                       breaks = c(90, 110)
                       ) +
  geom_sf(data = boundary, fill = NA,
          linewidth = 1) +
  labs(title = "Spatial-temporal Stability Map for standardized yield mean",
       fill = "Mean Pixel") +
  mytheme_map +
  annotation_scale(location = "bl",
                   width_hint = 0.3,
                   pad_x = unit(0.1, "in"),
                   pad_y = unit(0.01, "in")) +
  annotation_north_arrow(location = "tl",
                       pad_x = unit(0.1, "in"),
                       pad_y = unit(0.05, "in"),
                       height = unit(0.4, "in"),
                       width = unit(0.4, "in"),
                       style = north_arrow_fancy_orienteering())

sy_all_mean
```

Standardized  cv map  
```{r cv map}
sy_all_cv =  ggplot() +
  geom_sf(data = sy_all,
          aes(fill = cv_pixel),
          color = NA) +
  scale_fill_viridis_b(option = "C",
                       breaks = c(20)) +
                       #limit = c(0, 30)) 
   geom_sf(data = boundary, fill = NA,
          linewidth = 1) +
  labs(title = "Spatial-temporal Stability Map for standardized yield CV",
       fill = "CV Pixel") +
  mytheme_map 

sy_all_cv

```

```{r both maps}
sy_all_mean + sy_all_cv

ggsave("../output/sy-temp-stability.png",
       width = 12, height = 4)

```

# Classification  

Here's how I want to classify our spatial-temporal classes:  

- If a pixel has CV > 20%, then it is classified as "unstable"  

- If a pixel has CV < 20%, and mean standardized yield > 110%, then it is classified as "high-stable"  

- If a pixel has CV < 20%, and mean standardized yield > 90% and < 110%, then it is classified as "medium-stable" 

- If a pixel has CV < 20%, and mean standardized yield < 90%, then it is classified as "low-stable" 

Let's apply these rules below:  

```{r creating classes}
sy_all_class = sy_all %>%
  mutate(stclass = case_when(
    cv_pixel > 20 ~ "unstable",
    cv_pixel < 20 & mean_pixel >= 110 ~ "high-stable",
    cv_pixel < 20 & mean_pixel >= 90 & mean_pixel < 110 ~ "medium-stable",
    cv_pixel < 20 & mean_pixel < 90 ~ "low-stable"
    
  ))

sy_all_class 
```

Now, let's plot the classes.  
```{r plotting stability classes}
ggplot() +
  geom_sf(data = sy_all_class,
          aes(fill = stclass),
          color = NA) +
  scale_fill_viridis_d(option = "C") +
  geom_sf(data = boundary, fill = NA,
          linewidth = 1) +
  labs(title = "Yield Stability Map for 4 years for Mr McCormick",
       fill = "Stability Classes") +
  mytheme_map +
  annotation_scale(location = "bl",
                   width_hint = 0.3,
                   pad_x = unit(0.1, "in"),
                   pad_y = unit(0.01, "in")) +
  annotation_north_arrow(location = "tl",
                       pad_x = unit(0.1, "in"),
                       pad_y = unit(0.05, "in"),
                       height = unit(0.4, "in"),
                       width = unit(0.4, "in"),
                       style = north_arrow_fancy_orienteering())

ggsave("../output/yield-stability-map.png",
       width = 8, height = 5)
```



# Classes proportion  
 
```{r proportion per stability zone}
pix_prop = sy_all_class %>%
  group_by(stclass) %>%
  tally() %>%
  mutate(Proportion = n/6359 *100) %>%
  st_drop_geometry() %>%
  rename(`Observations (6369)` = n,
         `Stability-Classes` = stclass) %>%
  gt() %>%
  tab_header(title = "Proportion of Stability Classes") %>%
  fmt_number(columns = c(Proportion), decimals = 1) %>%
  fmt_number(columns = `Observations (6369)`, decimals = 0, use_seps = TRUE) %>%
  tab_options(
    table.width = pct(90),
    table.font.size = 14,
    data_row.padding = px(8),
    column_labels.font.weight = "bold",
    heading.title.font.size = 18,
    heading.title.font.weight = "bold"
  )

pix_prop
write_rds(pix_prop, "../output/stab-class_prop.rds")

```

# Exporting  
 
```{r exporting yield stability}
write_sf(sy_all_class, "../data/sy_all.geojson",
         delete_dsn = T)
```

