---
title: "Terrain properties"
author: "Comfort Adegbenro"
format:
  html:
    code-fold: false
    embed-resources: true
    toc: true
    number-sections: true
    theme: journal
    #cerulean, journal, quartz, sketchy, slate, morph, superhero, vapor
knitr: 
  opts_chunk: 
    message: false
    warning: false
---

# Objectives  
I will be importing only the corn cleaned yield data that contains the elevation data. This has more data point, and performing a summary for all other years show that thy re within range (difference of 12 ft, approximately). From this I will interpolate elevation, and then derive other terrain variables, make maps, and export all layers to file.


# Setup  
```{r loading libraries}

# Packages
library(dplyr)
library(tidyr)
library(janitor)
library(readr)
library(sf) #vector manipulation
library(ggplot2)
library(viridis)
library(ggthemes)
library(gstat) #IDW
library(stars) #raster manipulation
library(terra) #topographical variables
library(patchwork) #combining multiple plots  
library(tibble) #rownames to column

```

```{r defining map theme}
mytheme_map <- 
  theme_map()+
  theme(legend.position = "bottom",
        legend.direction = "vertical",
        panel.background = element_rect(fill="white",
                                        color=NA),
        plot.background = element_rect(fill="white",
                                       color=NA),
        legend.title = element_text(size=10),
        legend.text = element_text(size=9),
        plot.margin = margin(5, 10, 5, 10))

```

```{r reading elevation and boundary data and wrangling}
elev = read_sf("../data/y16_corn_clean.geojson") %>%
  mutate(elev_m = elev_ft * 0.3048) %>%
  dplyr::select(elev_m) 

elev
summary(elev)

boundary = read_sf("../data/50ac/boundary/BackFieldPivot.shp") %>%
  st_transform(6345)
```




# EDA  
```{r elev map}
elev %>%
  ggplot()+
  geom_sf(aes(color = elev_m), size =.01)+
  scale_color_viridis_c(option = "E") +
  geom_sf(data = boundary, fill = NA,
          linewidth = 1) +
  mytheme_map

```


# Grid

```{r grid_r}
# grid in raster format
grid_r = boundary %>%
  st_make_grid(cellsize = 5) %>% #5 means 5 meters because the boundary is in meters
  st_as_sf() %>%
  st_rasterize(dx = 5, dy = 5) %>%
  st_crop(boundary)

grid_r

ggplot() +
  geom_stars(data = grid_r)

```

# Interpolation  
```{r idw_mod}
idw_mod = gstat(formula = elev_m ~ 1,
                data = elev)

idw_mod
```

```{r idw_pred}
idw_pred = predict(idw_mod, grid_r)

idw_pred
```

```{r elev_map }
elev_map <- ggplot() +
  geom_stars(data = idw_pred) +
  geom_sf(data = boundary, fill = NA,
          linewidth = 1) +
  scale_fill_viridis_b(option = "E",
                       na.value = "white") +
  labs(title = "Elevation",
       fill = "Elev. (m)") +
  mytheme_map +
  annotation_scale(location = "bl",
                   width_hint = 0.3,
                   pad_x = unit(0.1, "in"),
                   pad_y = unit(0.01, "in")) +
  annotation_north_arrow(location = "tl",
                       pad_x = unit(0.1, "in"),
                       pad_y = unit(0.05, "in"),
                       height = unit(0.4, "in"),
                       width = unit(0.4, "in"),
                       style = north_arrow_fancy_orienteering())
  

elev_map
```

# Computing secondary variables  
## Slope  
```{r slope}
slope <- idw_pred %>%
  dplyr::select(elev_m = var1.pred) %>%
  
  # Transforming to SpatRast for slope computation  
  terra::rast() %>%
  terra::terrain(v = "slope",
                 unit = "radian") %>%

  # Transforming back into stars (so works with ggplot)
  st_as_stars() %>%

  # Transforming to pct
  mutate(pct = tan(slope)*100) %>%
  dplyr::select(pct)


slope
```


```{r slope map}
slope_map <- ggplot()+
  geom_stars(data = slope)+
  geom_sf(data = boundary, 
          fill = NA,
          linewidth = 1)+
  scale_fill_viridis_b(option = "E",
                       na.value = "white")+
  labs(title = "Slope",
       fill = "Slope\n(%)")+
  mytheme_map

slope_map
```

## Aspect  
```{r aspect}
aspect <- idw_pred %>%
  dplyr::select(elev_m = var1.pred) %>%
  terra::rast() %>%
  terra::terrain(v = "aspect",
                 unit = "degrees") %>%
  st_as_stars() %>%
  st_set_crs(value =  6345) # not compulsory to include this

aspect
```


```{r aspect map}
aspect_map <- ggplot()+
  geom_stars(data = aspect)+
  geom_sf(data = boundary, 
          fill = NA,
          linewidth = 1)+
  scale_fill_viridis_b(option = "E",
                       breaks = c(22.5,
                                  67.5,
                                  112.5,
                                  157.5,
                                  202.5,
                                  247.5,
                                  292.5,
                                  337.5),
                       na.value = "white")+
  labs(title = "Aspect",
       fill="Aspect\n(degrees)")+
  mytheme_map

aspect_map
```

## Flow direction  
```{r flowdir}
flowdir <- idw_pred %>%
  dplyr::select(elev_m = var1.pred) %>%
  terra::rast() %>%
  terra::terrain("flowdir",
                 unit = "radians") %>%
  st_as_stars() %>%
  st_set_crs(value =  6345)

flowdir
summary(flowdir)
summary(aspect)
summary(idw_pred)
```

The NAs are coming from the near-the-boundary cells, which were dropped off as the cells surrounding them are not available to compute the cell values.
```{r flowdir map}
flowdir_map <- ggplot()+
  geom_stars(data = flowdir)+
  geom_sf(data = boundary, 
          fill = NA,
          linewidth = 1)+
  scale_fill_viridis_b(option = "E",
                       breaks = c(1,2,
                                  4,8,
                                  16,32,
                                  64,128))+
  labs(title = "Flow Direction",
       fill="Flow dir.")+
  mytheme_map

flowdir_map
```

```{r flowdir map 2}
ggplot() +
  geom_stars(data = flowdir %>%
               mutate(flowdir = factor(flowdir)))+
  geom_sf(data = boundary, 
          fill = NA)+
  labs(fill="Flow dir.")+
  mytheme_map +
  facet_wrap(~flowdir) +
  theme(legend.position = "none")

```


## Plotting all together
```{r plotting all together}
elev_map + slope_map + aspect_map + flowdir_map +
  plot_layout(ncol = 4) +
  plot_annotation(title = "Interpolated Terrain Maps",
    theme = theme(
      plot.title = element_text(size = 13, face = "bold", hjust = 0.5)))

ggsave("../output/ip-terrain_maps.png",
       width = 18,
       height = 5)
```


# Grid extraction  
```{r elev_v}
elev_v = idw_pred %>%
  st_as_sf(as_points = F,
           merge = F) %>%
  dplyr::select(elev_m = var1.pred)

elev_v
```


```{r slope_v}
slope_v = slope %>%
  st_as_sf(as_points = F,
           merge = F) %>%
  dplyr::select(slope = pct)

slope_v
```
How many rows? Why is that?  We have lower rows because the cells near the boundary do not have neighbours so we cannot estimate the slope.

```{r aspect_v}
aspect_v <- aspect %>%
  st_as_sf(as_points = F, 
           merge = F)
  

aspect_v
```


```{r flowdir_v}
flowdir_v <- flowdir %>%
  st_as_sf(as_points = F, 
           merge = F)
flowdir_v
```

# Joining all topo variables in one object  

```{r terrain_v }
terrain_v = elev_v %>%
  st_join(slope_v,
          join = st_equals,
          left = T) %>%
  st_join(aspect_v,
          join = st_equals,
          left = T) %>%
  st_join(flowdir_v,
          join = st_equals,
          left = T)

terrain_v %>%
  summary()
```

# Exporting  
Exporting to file to use later for management zone creation.  
```{r exporting}
write_sf(terrain_v, 
         "../data/terrain_v.geojson",
         delete_dsn = T)

```





