---
author: "Comfort Adegbenro"
title: "Zone-based Variable rate nitrogen prescription"
output: html_document
---

#   Objectives  
The learning objectives for this exercise are to:  
  - Learn where to find crop-specific nutrient recommendations  
  - Understand what information is needed to recommend N in cotton   
  - Estimate zone-specific cotton yield potential  
  - Calculate zone-specific total N recommendation 
  - Calculate zone-specific side-dress variable N rate recommendation  
  - Calculate total amount of fertilizer to order  
  - Export recommendation to file to be used in a variable rate controller  

# Introduction  
We are going to create a recommendation to apply variable rate nitrogen (N) fertilizer in **corn**, with a **different in-season side-dress rate** for each zone.    

> How do we know how much N corn needs (or really, how do we know how much of any given nutrient any crop needs)?  

# Setup  
```{r setup}
# Packages
library(dplyr)
library(tidyr)
library(readr)
library(sf) #vector manipulation
library(ggplot2)
library(viridis)
library(ggthemes)

```

```{r defining map theme}
mytheme_map <- 
  theme_map()+
  theme(legend.position = "bottom",
        legend.direction = "vertical",
        panel.background = element_rect(fill="white",
                                        color=NA),
        plot.background = element_rect(fill="white",
                                       color=NA),
        legend.title = element_text(size=10),
        legend.text = element_text(size=9),
        plot.margin = margin(5, 10, 5, 10))

```


```{r y16_corn}
y16_ip_corn = read_sf("../data/y16_corn_ip.geojson") 
  
y16_ip_corn

summary(y16_ip_corn)

boundary = read_sf("../data/50ac/boundary/BackFieldPivot.shp") %>%
  st_transform(crs = 6345)
```


```{r zone_st}
zone_st = read_sf("../data/zone_s.geojson") %>%
  dplyr::select(zone, geometry)

summary(zone_st)
```

# Wrangling  
```{r Merging yield and zones}
zone_yield <- zone_st %>%
  st_join(y16_ip_corn,
          join = st_nearest_feature, 
          left = T
  )

summary(zone_yield)

ggplot() +
  geom_sf(data = zone_yield,
          aes(fill = zone))

```

# VRN  
Below, we will
- Calculate zone-specific yield potential (as the 90% quantile of the yield distribution of only the 2016 corn year)  

- Calculate the total N rate (lbs N/ac) needed with **UGFertex**
- Adjust it in +25% assuming that we are following cotton  
- Adjust it in 2/3 to be applied side-dress (1/3 applied at planting)  
- Define our N source as urea-ammonium nitrate (UAN) 28% N  
- Calculate the total product rate (gal UAN/ac)  

```{r zone yield potential}
zone_yield %>%
  ggplot(aes(x = corn_ipyield_buac, fill = zone))+
  geom_density(alpha = .6) 

```

```{r yield potential by zone}
yp_zone = zone_yield %>%
  group_by(zone) %>%
  summarize(yieldpotential = quantile(corn_ipyield_buac, 0.9, na.rm = TRUE)) %>%
  st_drop_geometry()
write_csv(yp_zone, "../output/yp_zone.csv")

yp_zone

```


```{r vrn calculation }
vrn = zone_yield %>%
  group_by(zone) %>%
  summarize(yieldpotential = quantile(corn_ipyield_buac, 0.9, na.rm = TRUE)) %>%
  
    # Stop here, include these on plot above

  # Go on UGFertex and enter these yield potential values, assuming cotton as previous crop, irrigated, coastal plain soil. What total N did you get for zone high yield? What about zone low yield?
  
  # We can also calculate those using the formula below, which is the formula used behind the scenes on UGFertex:
  mutate(totalnrate_lbsac = yieldpotential*1.2) %>% 
 # Calculate pre-plant rate
  mutate(pp_totalnrate_adj_lbsac = totalnrate_lbsac * 0.30) %>%
  # Calculate ~30% as portion to be applied in-season (Assuming a single pre-plant rate was applied to both zones)
  mutate(is_totalnrate_lbsac = round(totalnrate_lbsac - 72.5, 1)) %>%
  # Calculate product rate when using UAN solution 28% N
  # UAN 28% has 3 lbs N per gallon of product (see chunk below)
  mutate(uanrate_galac = round(is_totalnrate_lbsac/3, 1))

vrn

write_csv(vrn %>%
            st_drop_geometry(), "../output/total_nrate.csv")

```
UAN is a liquid fertilizer, so how do we go from volume of solution to lbs of N?  

> What information do we need?  










We can find this information on the fertilizer label/msds information from manufacturer.  

Normally, UAN 28% density is 10.67 lbs/gal, as we can find from these retailers:  
- https://www.agricocanada.com/wp-content/uploads/2016/09/agric-101-SDS_Urea-Ammonium-Nitrate-UAN-Solution.pdf  
-  https://www.cfindustries.com/globalassets/cf-industries/media/documents/safety-data-sheets/uan---north-america/urea_ammonium_nitrate_solution_uan_sds.pdf  

NOTE: UAN is normally sold as 28, 30, or 32%N. Each of them has a different density!  


```{r UAN volume to mass of N}
uan28density_lbsgal <- 10.67

uan28Nconc_pct <- 0.28

uan28_lbsNgal <- uan28density_lbsgal * uan28Nconc_pct

uan28_lbsNgal #~3 lbs N/gal of UAN28%

```

How much UAN28% do we need to have delivered at the field to ensure enough product for application?  

```{r total volume of UAN per area}
vrn %>%
  mutate(area_m2 = st_area(.)) %>%
  mutate(area_ac = as.numeric(area_m2/4047)) %>%

  # Will ask for 20% more than recommended to have a bit extra 
  mutate(uan_gal = (uanrate_galac*area_ac)*1.2) %>%
  summarize(totaluan_gal = sum(uan_gal))

```

# Rx map  
```{r vrn vector}
vrn_v = zone_yield %>%
  left_join(vrn %>%
              st_drop_geometry())

vrn_v
```

```{r in-season N rate vrn plot}
vrn_v %>%
  mutate(is_totalnrate_lbsac = factor(is_totalnrate_lbsac)) %>%
  ggplot()+
  geom_sf(aes(fill = is_totalnrate_lbsac,
              color = is_totalnrate_lbsac)) +
  scale_fill_colorblind() +
  scale_color_colorblind() +
  geom_sf(data = boundary, fill = NA,
          linewidth = 1) +
  labs(title="VRN Rx - In-season N rate",
       fill="N rate\n(lb N/ac)",
       color="N rate\n(lb N/ac)")+
  mytheme_map +
  annotation_scale(location = "bl",
                   width_hint = 0.3,
                   pad_x = unit(0.1, "in"),
                   pad_y = unit(0.01, "in")) +
  annotation_north_arrow(location = "tl",
                       pad_x = unit(0.1, "in"),
                       pad_y = unit(0.05, "in"),
                       height = unit(0.4, "in"),
                       width = unit(0.4, "in"),
                       style = north_arrow_fancy_orienteering()) +
  theme(plot.title = element_text(size = 13, face = "bold", hjust = 0.5))


ggsave("../output/inseason-Nrate-vrn.png",
      width = 6, 
       height = 4
       )  
```


```{r UAN vrn plot}
vrn_v %>%
  mutate(uanrate_galac=factor(uanrate_galac)) %>%
  ggplot()+
  geom_sf(aes(fill = uanrate_galac,
              color = uanrate_galac)) +
  geom_sf(data = boundary, fill = NA,
          linewidth = 1) +
  labs(title="VRN Rx - UAN 28% rate",
       fill="UAN rate\n(gal/ac)",
       color="UAN rate\n(gal/ac)")+
  mytheme_map +
  annotation_scale(location = "bl",
                   width_hint = 0.3,
                   pad_x = unit(0.1, "in"),
                   pad_y = unit(0.01, "in")) +
  annotation_north_arrow(location = "tl",
                       pad_x = unit(0.1, "in"),
                       pad_y = unit(0.05, "in"),
                       height = unit(0.4, "in"),
                       width = unit(0.4, "in"),
                       style = north_arrow_fancy_orienteering())+
  theme(plot.title = element_text(size = 13, face = "bold", hjust = 0.5))



ggsave("../output/UAN-vrn.png",
      width = 6, 
       height = 4
       )  
```

# Exporting  
We are going to export as **shapefile** as it is the most commonly accepted format in variable rate controllers.  

Since we are exporting as shapefile, let's create a sub-folder inside `output` called `VRN` to have all the associated files saved together.  

```{r exporting}
vrn_v %>%
  dplyr::select(zone, is_totalnrate_lbsac, uanrate_galac) %>%
  write_sf("../output/vrn.shp",
           delete_dsn = T)
```

# Summary  
In summary, today we:  
- Understood where to look for information on crop and nutrient-specific recommendations  
- Defined yield potential for each zone  
- Calculated VRN in both lbs N/ac and gallons UAN38/ac  
- Exported VRN Rx as a shapefile  





