---
title: "Satellite mean of all 4 years and spatial temporal stability"
format: html
---

# Learning objectives

In this exercise, our goals are to:\
- Retrieve the median GNDVI across the growing season for 3 separate years\
- Conduct spatial-temporal stability analysis using the median GNDVI (instead of yield)\
- Compare GNDVI-derived spatial-temporal classes with the one based on yield.

# Introduction

Performing spatial-temporal yield stability analysis is an important step in precision agriculture as it sets the stage for yield expectation and patterns in space and time.

This information can then be used to address yield limiting factors and serve as basis for variable rate management of certain inputs.

Now, what if:\
- the grower does not have yield monitoring capacity?\
- the crop being grown does not have yield monitoring technology developed?

Perhaps we can use satellite imagery to create the zones instead.

Let's give it a try.

# Setup

```{r}
library(openeo)
library(mapview)
library(tidyverse)
library(stars)
library(sf)
library(mapview)
library(ggthemes)

```



```{r defining map theme}
mytheme_map <- 
  theme_map()+
  theme(legend.position = "bottom",
        legend.direction = "vertical",
        panel.background = element_rect(fill="white",
                                        color=NA),
        plot.background = element_rect(fill="white",
                                       color=NA),
        legend.title = element_text(size=10),
        legend.text = element_text(size=9),
        plot.margin = margin(5, 10, 5, 10))

```


Importing field boundary


```{r}
field <- read_sf("../data/50ac/boundary/BackFieldPivot.shp") %>%
  st_transform(crs = 6345) %>%
  st_buffer(dist = -10) # because values close to and outside the border are normally higher and thus skew color scales when plotting

field

boundary <- read_sf("../data/50ac/boundary/BackFieldPivot.shp") %>%
  st_transform(crs = 6345) 
```

```{r}
mapview(field)
```

# Connecting to the API

```{r}
connection <- connect(host = "https://openeo.dataspace.copernicus.eu")
```

# Login credentials

```{r}
login() #after running this, go to console and hit ENTER/RETURN on your keyboard to be prompted to browser to login.  
```

# Starting a process

```{r}
p <- processes()
```

# Imagery for 2016

## Setting data retrieval parameters

Here we need to determine our query parameters, including:

-   **Collection and data product** we want to pull from

```{r}
data_16 <- p$load_collection(id = "SENTINEL2_L2A", 
                         spatial_extent = field,
                         temporal_extent = list(
                           "2016-05-01", # start date
                           "2016-10-15" # end date
                           ),
                         bands = list("B03","B08", "SCL") #B03 = green, B08 = NIR, SCL contains land classification used to remove clouds
)

data_16
```

## Cloud masking

```{r}
# define filter function to create mask
filter_function <- function(data, context) {
  vegetation <- p$eq(data["SCL"], 4) # vegetation is 4
  non_vegetation <- p$eq(data["SCL"], 5) # non-vegetation is 5
  # we want to mask all other values, so NOT (4 OR 5)
  return(p$not(p$or(vegetation, non_vegetation)))
}

# create mask by reducing bands with our defined formula
cloud_reduce <- p$reduce_dimension(data = data_16, 
                                   reducer = filter_function, 
                                   dimension = "bands")

# mask the bands data
cloud_reduced <- p$mask(data_16, cloud_reduce)
```

## Spectral reducer (GNDVI)

```{r}

spectral_reduce <- p$reduce_dimension(data = cloud_reduced, 
                                      dimension = "bands",
                                      reducer = function(data,context) {
                                        b3 = data["B03"]
                                        b8 = data["B08"]
                                        
                                        return((b8-b3)/(b8+b3))
                                      })
```

## Temporal reducer (mean)

```{r}
temporal_reduce <- p$reduce_dimension(data = spectral_reduce,
                                      dimension = "t", 
                                      reducer = function(x,y){
                                        mean(x, na.rm = T)
                                        
                                      })



```

## Compiling result

```{r}
result <- p$save_result(data=temporal_reduce,
                        format= "GTiff") 

```

Up until the chunk above, these all are sent to the server for processing.

Once the processing is done on the server, we download the result (next chunk).

## Downloading result

```{r}
compute_result(graph = result, 
               output_file = "../output/2016_mean_GNDVI.gtif")
```

## Importing result to check

```{r}
gndvi_mean16 <- read_stars("../output/2016_mean_GNDVI.gtif") %>%
  st_transform(crs = 6345) %>%
  rename(gndvi = 1) %>%
  st_crop(field)

gndvi_mean16

```

```{r}
mapview(gndvi_mean16) # this map is the product of getting the median of each pixel over an entire growing season, so it does not belong to any one date. So a large median of gndvi would be assumed to give more yield, but high seedling vigor does not mean it will translate to yield for cotton. But for corn, soybean, the more vegetative biomass you have, the more yield. For cotton, it could be a tradeoff between reproductive and vegetative, because it is perennial.
```

```{r gndvi_16_map}
gndvi_16_map <- ggplot()+
  geom_stars(data = gndvi_mean16,
             color = NA) +
  theme_map() +
  labs(title = "Corn (2016)",
       fill = "GNDVI") +
  scale_fill_viridis_c() +
  geom_sf(data = boundary, fill = NA,
          linewidth = 1) +
  mytheme_map +
  annotation_scale(location = "bl",
                   width_hint = 0.3,
                   pad_x = unit(0.1, "in"),
                   pad_y = unit(0.01, "in")) +
  annotation_north_arrow(location = "tl",
                       pad_x = unit(0.1, "in"),
                       pad_y = unit(0.05, "in"),
                       height = unit(0.4, "in"),
                       width = unit(0.4, "in"),
                       style = north_arrow_fancy_orienteering())
  

gndvi_16_map
```

# Imagery for 2017  

Now, let's pull GNDVI mean for 2017.

## Setting data retrieval parameters  

Here we need to determine our query parameters, including:

-   **Collection and data product** we want to pull from

```{r}
data_17 <- p$load_collection(id = "SENTINEL2_L2A", 
                         spatial_extent = field,
                         temporal_extent = list(
                           "2017-05-01", # start date
                           "2017-10-15" # end date
                           ),
                         bands = list("B03","B08", "SCL") #B03 = green, B08 = NIR, SCL contains land classification used to remove clouds
)

data_17
```

## Cloud masking

```{r}
# create mask by reducing bands with our defined formula
cloud_reduce_17 <- p$reduce_dimension(data = data_17, 
                                   reducer = filter_function, 
                                   dimension = "bands")

# mask the bands data
cloud_reduced_17 <- p$mask(data_17, cloud_reduce_17)
```

## Spectral reducer (GNDVI)

```{r}

spectral_reduce_17 <- p$reduce_dimension(data = cloud_reduced_17, 
                                      dimension = "bands",
                                      reducer = function(data,context) {
                                        b3 = data["B03"]
                                        b8 = data["B08"]
                                        
                                        return((b8-b3)/(b8+b3))
                                      })
```

## Temporal reducer (mean)

```{r}
temporal_reduce_17 <- p$reduce_dimension(data = spectral_reduce_17,
                                      dimension = "t", 
                                      reducer = function(x,y){
                                        mean(x, na.rm = T)
                                      })


```

## Compiling result

```{r}
result_17 <- p$save_result(data=temporal_reduce_17,
                        format= "GTiff") 

```

Up until the chunk above, these all are sent to the server for processing.

Once the processing is done on the server, we download the result (next chunk).

## Downloading result

```{r}
compute_result(graph = result_17, 
               output_file = "../output/2017_mean_GNDVI.gtif")
```

## Importing result to check

```{r}
gndvi_mean_17 <- read_stars("../output/2017_mean_GNDVI.gtif") %>%
  st_transform(crs = 6345) %>%
  rename(gndvi = 1) %>%
  st_crop(field)

gndvi_mean_17

```

```{r}
mapview(gndvi_mean_17)
```

```{r gndvi_17_map}
gndvi_17_map <- ggplot()+
  geom_stars(data = gndvi_mean_17,
             color = NA) +
  labs(title = "Soybeans (2017)") +
  scale_fill_viridis_c() +
  geom_sf(data = boundary, fill = NA,
          linewidth = 1) +
  mytheme_map


gndvi_17_map

```

# Imagery for 2023  

Now, let's pull GNDVI mean for 2023.

## Setting data retrieval parameters

Here we need to determine our query parameters, including:

-   **Collection and data product** we want to pull from

```{r}
data_23 <- p$load_collection(id = "SENTINEL2_L2A", 
                         spatial_extent = field,
                         temporal_extent = list(
                           "2023-06-01", # start date
                           "2023-11-30" # end date
                           ),
                         bands = list("B03","B08", "SCL") #B03 = green, B08 = NIR, SCL contains land classification used to remove clouds
)

data_23
```

## Cloud masking

```{r}
# create mask by reducing bands with our defined formula
cloud_reduce_23 <- p$reduce_dimension(data = data_23, 
                                   reducer = filter_function, 
                                   dimension = "bands")

# mask the bands data
cloud_reduced_23 <- p$mask(data_23, cloud_reduce_23)
```

## Spectral reducer (GNDVI)

```{r}

spectral_reduce_23 <- p$reduce_dimension(data = cloud_reduced_23, 
                                      dimension = "bands",
                                      reducer = function(data,context) {
                                        b3 = data["B03"]
                                        b8 = data["B08"]
                                        
                                        return((b8-b3)/(b8+b3))
                                      })
```

## Temporal reducer (mean)

```{r}
temporal_reduce_23 <- p$reduce_dimension(data = spectral_reduce_23,
                                      dimension = "t", 
                                      reducer = function(x,y){
                                        mean(x, na.rm = T)
                                      })


```

## Compiling result

```{r}
result_23 <- p$save_result(data=temporal_reduce_23,
                        format= "GTiff") 

```

Up until the chunk above, these all are sent to the server for processing.

Once the processing is done on the server, we download the result (next chunk).

## Downloading result

```{r}
compute_result(graph = result_23, 
               output_file = "../output/2023_mean_GNDVI.gtif")
```

## Importing result to check

```{r}
gndvi_mean_23 <- read_stars("../output/2023_mean_GNDVI.gtif") %>%
  st_transform(crs = 6345) %>%
  rename(gndvi = 1) %>%
  st_crop(field)

gndvi_mean_23

```

```{r}
mapview(gndvi_mean_23)
```

```{r gndvi_23_map}
gndvi_23_map <- ggplot()+
  geom_stars(data = gndvi_mean_23,
             color = NA) +
  theme_map() +
  labs(title = "Cotton (2023)")+
  scale_fill_viridis_c() +
  geom_sf(data = boundary, fill = NA,
          linewidth = 1) +
  mytheme_map


gndvi_23_map
```


# Imagery for 2024  

Now, let's pull GNDVI mean for 2024.

## Setting data retrieval parameters

Here we need to determine our query parameters, including:

-   **Collection and data product** we want to pull from

```{r}
data_24 <- p$load_collection(id = "SENTINEL2_L2A", 
                         spatial_extent = field,
                         temporal_extent = list(
                           "2024-06-01", # start date
                           "2024-11-30" # end date
                           ),
                         bands = list("B03","B08", "SCL") #B03 = green, B08 = NIR, SCL contains land classification used to remove clouds
)

data_24
```

## Cloud masking

```{r}
# create mask by reducing bands with our defined formula
cloud_reduce_24 <- p$reduce_dimension(data = data_24, 
                                   reducer = filter_function, 
                                   dimension = "bands")

# mask the bands data
cloud_reduced_24 <- p$mask(data_24, cloud_reduce_24)
```

## Spectral reducer (GNDVI)

```{r}

spectral_reduce_24 <- p$reduce_dimension(data = cloud_reduced_24, 
                                      dimension = "bands",
                                      reducer = function(data,context) {
                                        b3 = data["B03"]
                                        b8 = data["B08"]
                                        
                                        return((b8-b3)/(b8+b3))
                                      })
```

## Temporal reducer (median)

```{r}
temporal_reduce_24 <- p$reduce_dimension(data = spectral_reduce_24,
                                      dimension = "t", 
                                      reducer = function(x,y){
                                        mean(x, na.rm = T)
                                      })


```

## Compiling result

```{r}
result_24 <- p$save_result(data=temporal_reduce_24,
                        format= "GTiff") 

```

Up until the chunk above, these all are sent to the server for processing.

Once the processing is done on the server, we download the result (next chunk).

## Downloading result

```{r}
compute_result(graph = result_24, 
               output_file = "../output/2024_mean_GNDVI.gtif")
```

## Importing result to check

```{r}
gndvi_mean_24 <- read_stars("../output/2024_mean_GNDVI.gtif") %>%
  st_transform(crs = 6345) %>%
  rename(gndvi = 1) %>%
  st_crop(field)

gndvi_mean_24

```

```{r}
mapview(gndvi_mean_24)
```

```{r gndvi_24_map}
gndvi_24_map <- ggplot()+
  geom_stars(data = gndvi_mean_24,
             color = NA) +
  labs(title = "Cotton (2024)")+
  scale_fill_viridis_c() +
  geom_sf(data = boundary, fill = NA,
          linewidth = 1) +
  mytheme_map


gndvi_24_map
```


# Combining all maps

```{r}
library(patchwork)
gndvi_16_map + gndvi_17_map + gndvi_23_map + gndvi_24_map +
  plot_layout(ncol = 4)
  
```

# Spatial-temporal stability analysis

## From raster to vector

First, let's transform from raster to vector. For 2016:

```{r}
gndvi_mean16_v <- gndvi_mean16 %>%
  st_as_sf() %>%
  rename(gndvi_16 = gndvi)

gndvi_mean16_v
```

For 2019:

```{r}
gndvi_mean_17_v <- gndvi_mean_17 %>%
  st_as_sf() %>%
  rename(gndvi_17 = gndvi)

gndvi_mean_17_v

```

For 2023:

```{r}
gndvi_mean_23_v <- gndvi_mean_23 %>%
  st_as_sf() %>%
  rename(gndvi_23 = gndvi)

gndvi_mean_23_v

```


For 2024:

```{r}
gndvi_mean_24_v <- gndvi_mean_24 %>%
  st_as_sf() %>%
  rename(gndvi_24 = gndvi)

gndvi_mean_24_v

```

# Standardizing with the field-level median

Beginning with 2017 yield data.

```{r sy_17}
sy_16 = gndvi_mean16_v %>%
  mutate(median_gndvi = median(gndvi_16)) %>%
  mutate(sy = gndvi_16/median_gndvi*100) %>%
  mutate(sy = round(sy, 1)) %>%
  dplyr::select(sy16 = sy, geometry)
  

sy_16
summary(sy_16)
```

Let's check how does this standardization looks like in space.

```{r sy_16 map}
sy_16_map <- ggplot() +
  geom_sf(data = sy_16,
          aes(fill = sy16),
          color = NA
          ) +
  scale_fill_viridis_c(breaks = seq(80, 110, 5),
                       limits = c(80, 110)) +
  geom_sf(data = boundary, fill = NA,
          linewidth = 1) +
  labs(title = "2016",
       fill = "Median-\nstandardized \nGNDVI") +
  mytheme_map +
  annotation_scale(location = "bl",
                   width_hint = 0.3,
                   pad_x = unit(0.1, "in"),
                   pad_y = unit(0.01, "in")) +
  annotation_north_arrow(location = "tl",
                       pad_x = unit(0.1, "in"),
                       pad_y = unit(0.05, "in"),
                       height = unit(0.4, "in"),
                       width = unit(0.4, "in"),
                       style = north_arrow_fancy_orienteering())


sy_16_map 
```

Repeat the same steps for 2017.

```{r sy_17}
sy_17 <- gndvi_mean_17_v %>%
  mutate(median_gndvi = median(gndvi_17)) %>%
  mutate(sy = gndvi_17/median_gndvi*100,
         sy = round(sy, 1)
         ) %>%
  dplyr::select(sy17 = sy, geometry)

sy_17

summary(sy_17)
```

```{r sy_17 map}
sy_17_map <- ggplot()+
  geom_sf(data = sy_17,
          aes(fill = sy17), 
          color = NA)+
  scale_fill_viridis_c(breaks = seq(80, 110, 5),
                       limits = c(80, 110)
  )+
  geom_sf(data = boundary, fill = NA,
          linewidth = 1) +
  labs(title="2017")+
  mytheme_map +
  theme(legend.position = "none")


sy_17_map 
```

Repeat the same steps for 2023.

```{r sy_23}
sy_23  <- gndvi_mean_23_v %>%
  mutate(median_gndvi = median(gndvi_23)) %>%
  mutate(sy = gndvi_23/median_gndvi*100,
         sy = round(sy, 1)
         ) %>%
  dplyr::select(sy23 = sy, geometry)

sy_23

summary(sy_23)
```

```{r sy_23 map}
sy_23_map <- ggplot()+
  geom_sf(data = sy_23,
          aes(fill = sy23), 
          color = NA)+
   scale_fill_viridis_c(breaks = seq(80, 110, 5),
                       limits = c(80, 110)
  )+
  geom_sf(data = boundary, fill = NA,
          linewidth = 1) +
  labs(title = "2023")+
  mytheme_map +
  theme(legend.position = "none")

sy_23_map 
```
Repeat the same steps for 2024.

```{r sy_24}
sy_24  <- gndvi_mean_24_v %>%
  mutate(median_gndvi = median(gndvi_24)) %>%
  mutate(sy = gndvi_24/median_gndvi*100,
         sy = round(sy, 1)
         ) %>%
  dplyr::select(sy24 = sy, geometry)

sy_24

summary(sy_24)
```

```{r sy_24 map}
sy_24_map <- ggplot()+
  geom_sf(data = sy_24,
          aes(fill = sy24), 
          color = NA)+
   scale_fill_viridis_c(breaks = seq(80, 110, 5),
                       limits = c(80, 110)
  )+
  geom_sf(data = boundary, fill = NA,
          linewidth = 1) +
  labs(title = "2024")+
  mytheme_map +
  theme(legend.position = "none")

sy_24_map 
```


Let's plot all 4 maps side-by-side to assist in visualizing them.

```{r all maps}
library(patchwork)

sy_16_map + sy_17_map + sy_23_map + sy_24_map +
  plot_layout(ncol = 4) +
  plot_annotation(title = "Median Standardized GNDVI Maps of Spatial Stability",
    theme = theme(
      plot.title = element_text(size = 13, face = "bold", hjust = 0.5)))

ggsave("../output/spatialstability_gndvi.png",
       width = 18, height = 5)

```

```{r}
summary(sy_16$sy16) #88 to 104
summary(sy_17$sy17) #81 to 107
summary(sy_23$sy23) #92 to 105
summary(sy_24$sy24) #92 to 106
```

# Spatio-temporal stability  

To perform a temporal analysis, we need to merge all FOUR data frames into one.

After joining the data sets, we still only have their information on year-specific yield spatial standardization.

> How can we merge these pieces of information to create a spatial-temporal classification?

We can use the yield mean and coefficient of variation (CV) of each pixel over 4 years, and use that to guide our decision.

First, let's merge the 4 data sets and calculate pixel-based mean and cv.

## Pixel-based mean and cv

```{r}
sy_all <- sy_16 %>%
  st_join(sy_17,
          join = st_equals,
          left = T
          ) %>%
  st_join(sy_23,
          join = st_equals,
          left = T
          ) %>%
   st_join(sy_24,
          join = st_equals,
          left = T
          ) %>%
  rowwise() %>%
  mutate(mean_pixel = mean(c(sy16, sy17, sy23, sy24))) %>%
  mutate(sd_pixel = sd(c(sy16, sy17, sy23, sy24))) %>%
  mutate(cv_pixel = (sd_pixel/mean_pixel)*100)

sy_all
summary(sy_all)

```

## EDA on pixel-based stats

Summary:

```{r summary mean }
summary(sy_all$mean_pixel)
```

Density plot:

```{r density plot mean }
ggplot(data = sy_all,
       aes(x = mean_pixel)
       ) +
  geom_density() +
  geom_rug()
  

```

```{r summary }
summary(sy_all$cv_pixel)
```

Density plot:

```{r density plot }
ggplot(data = sy_all,
       aes(x = cv_pixel)
       ) +
  geom_density() +
  geom_rug()
  

```

Standardized mean map

```{r mean map}
sy_all_mean = ggplot() +
  geom_sf(data = sy_all,
          aes(fill = mean_pixel),
          color = NA) +
  scale_fill_viridis_b(option = "C",
                       breaks = c(92, 102)
                       ) +
  geom_sf(data = boundary, fill = NA,
          linewidth = 1) +
  labs(title = "Spatial-temporal Stability Map for standardized GNDVI mean",
       fill = "Mean Pixel") +
  mytheme_map +
  annotation_scale(location = "bl",
                   width_hint = 0.3,
                   pad_x = unit(0.1, "in"),
                   pad_y = unit(0.01, "in")) +
  annotation_north_arrow(location = "tl",
                       pad_x = unit(0.1, "in"),
                       pad_y = unit(0.05, "in"),
                       height = unit(0.4, "in"),
                       width = unit(0.4, "in"),
                       style = north_arrow_fancy_orienteering())

sy_all_mean
```

Standardized mean cv map

```{r cv map}
sy_all_cv =  ggplot() +
  geom_sf(data = sy_all,
          aes(fill = cv_pixel),
          color = NA) +
  scale_fill_viridis_b(option = "C",
                       breaks = 5) +
                       #limit = c(0, 30)) 
   geom_sf(data = boundary, fill = NA,
          linewidth = 1) +
  labs(title = "Spatial-temporal Stability Map for standardized GNDVI CV",
       fill = "CV Pixel") +
  mytheme_map 

sy_all_cv

```

```{r both maps}
sy_all_mean + sy_all_cv

ggsave("../output/GNDVI_sy-temp-stability.png",
       width = 12, height = 4)

```

# Classification  

Here's how I want to classify our spatial-temporal classes:

-   If a pixel has CV \> 5%, then it is classified as "unstable"

-   If a pixel has CV \< 5%, and mean standardized yield \> 102%, then it is classified as "high-stable"

-   If a pixel has CV \< 5%, and mean standardized yield \> 92% and \< 102%, then it is classified as "medium-stable"

-   If a pixel has CV \< 5%, and mean standardized yield \< 92%, then it is classified as "low-stable"

> Given our summary for mean_pixel and cv_pixel, do you think these thresholds would work?

Let's apply these rules below:

```{r}
sy_all_class <- sy_all %>%
  mutate(stclass = case_when(
    cv_pixel > 5 ~ "unstable",
    cv_pixel < 5 & mean_pixel >= 102 ~ "high-stable",
    cv_pixel < 5 & mean_pixel < 102 & mean_pixel >= 92 ~ "medium-stable",
    cv_pixel < 5 & mean_pixel < 92 ~ "low-stable"
  )) %>%
  mutate(stclass = factor(stclass,
                          levels = c("high-stable",
                                     "medium-stable",
                                     "low-stable",
                                     "unstable")))

sy_all_class 
```

Now, let's plot the classes.

```{r stability classes}
ggplot() +
  geom_sf(data = sy_all_class,
          aes(fill = stclass),
          color = NA) +
  scale_fill_manual(
    values = c("high-stable" = "#0D0887",      # dark blue/purple
               "low-stable" = "#9C179E",        # purple/magenta
               "medium-stable" = "#ED7953",     # salmon/orange
               "unstable" = "yellow"),         # yellow
    drop = FALSE
  ) +
  geom_sf(data = boundary, fill = NA,
          linewidth = 1) +
  labs(title = "GNDVI Stability Map for 4 years for Mr McCormick",
       fill = "Stability Classes") +
  mytheme_map +
  annotation_scale(location = "bl",
                   width_hint = 0.3,
                   pad_x = unit(0.1, "in"),
                   pad_y = unit(0.01, "in")) +
  annotation_north_arrow(location = "tl",
                       pad_x = unit(0.1, "in"),
                       pad_y = unit(0.05, "in"),
                       height = unit(0.4, "in"),
                       width = unit(0.4, "in"),
                       style = north_arrow_fancy_orienteering())

ggsave("../output/gndvi-stclass-map.png",
       width = 8, height = 5)

```






